---
title: "TBI"
subtitle: "SMM069 (2025-26) Group Coursework - Group 12"
author:
  - "Gaurav Agrawal"
  - "Benjamin Evans"
  - "Samyak Jain"
  - "Basmah Khan"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    self_contained: true
    css: style.css
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: united
    highlight: tango
    df_print: paged
  # runtime: shiny
  bookdown::pdf_document2:
    includes:
      in_header: preamble.tex
    pandoc_args:
      - "--lua-filter=answer.lua"
    number_sections: true
    toc: true
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
# fontsize: 10pt
bibliography: references.bib
---

```{=latex}
\newpage
```

```{=html}
<h1>Information</h1>

This R Markdown document was created as part of a group assignment for SMM069 at Bayes Business School, City St George's, University of London in Term 3 2025-26.

```

```{r asthetic-header, include=FALSE, echo=FALSE}
# ============================================================================ #
# Key Information ====
# SMM069 Advanced Predictive Analytics
# Group Coursework 2025-26
# Group:        Group 12
# Authors (in alphabetical order):
#   - "Gaurav Agrawal"
#   - "Benjamin Evans"
#   - "Samyak Jain"
#   - "Basmah Khan"
# Professor:    Prof Vali Asimit
# Institution:  Bayes Business School - City St George's, University of London
# Date:         09 Jul 2026
#
# Description:  Term 3 group project for SMM069 Advanced Predictive Analtyics
#               (100% of coursework grade - 100% of module grade). The R code
#               below has been exported directly from an R Markdown (.rmd) file.
#               Hence the knitr settings.
#
# Dependencies (!TBI):
#   - tidyverse: loads dplyr, tidyr, ggplot2, purrr (& others)
#   - scales: to express portfolio weights as percentage
#   - patchwork: for combining multiple plots into one figure
#
# Acknowledgements
#   - TBI
# ============================================================================ #
```

```{r setup-knitr, include=FALSE, echo=FALSE}
# Setup & configuration ====
#----------------------- Initial setup (knitr settings) -----------------------#
dir.create("fig", showWarnings = FALSE)

# Defaults common to all outputs
knitr::opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  warning  = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.path  = "fig/",
  dpi       = 300
)

# Output-specific settings
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "pdf",
    fig.pos = "!ht",
    out.extra = ""
  )
} else {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "svglite"  # or "png"
  )
}
```

```{r setup-qol, include=FALSE, echo=FALSE}
#----------------------------- Clean environment ------------------------------#
rm(list = ls()) # Remove all objects
graphics.off() # Close all graphical devices
cat("\014") # Clean console
```

```{r load-dependencies, include=FALSE, echo=FALSE}
#------------------- Load dependencies / external libraries -------------------#
# --- Data Manipulation & Plotting ---
library(tidyverse) # loads dplyr, tidyr, ggplot2, purrr (& others)
library(patchwork) # for combining multiple plots into one figure

# --- Reporting & Tables ---
library(kableExtra) # for general tables
library(DT) # for HTML interactive data tables
library(plotly) # for interactive html plots
library(pander) # for writing to pandoc
```

```{r html-app, echo=FALSE, results='asis', eval=knitr::is_html_output(), purl=FALSE}
#------------------------------ HTML link to app ------------------------------#
library(htmltools)

# div(style = "background-color: #f8f9fa; padding: 20px; border: 1px solid #e9ecef; border-radius: 5px; text-align: center; margin-bottom: 30px;",
#   h3("Bonus Interactive Dashboard Available"),
#   p("This static report is accompanied by a live R Shiny dashboard allowing one to use a selection of 16 US stocks in different 5 portfolio combinations."),
#   a(href = "https://3enji.shinyapps.io/smm068-a01-g08-dashboard/",
#     target = "_blank",
#     class = "btn btn-primary",
#     style = "background-color: #007bff; color: white; padding: 10px 20px; text-decoration: none; border-radius: 5px; font-weight: bold;",
#     "Launch Interactive Dashboard")
# )
```

```{r custom-functions, include=FALSE, echo=FALSE}
#---------------------------- Custom QOL functions ----------------------------#
#####################################
# function: banner comments (used to to section up code)
# Usage: banner_comment("Element 1: data cleaning") -> then ctrl + v (or cmd+v)
#####################################
banner_comment <- function(text, width = 80, border = "#", fill = "-") {
  txt <- paste0(" ", text, " ")
  inner_width <- width - 2 * nchar(border)
  banner_string <- ""

  if (inner_width <= nchar(txt)) {
    banner_string <- paste0(border, txt, border)
  } else {
    pad_total <- inner_width - nchar(txt)
    pad_left <- pad_total %/% 2
    pad_right <- pad_total - pad_left

    banner_string <- paste0(
      border,
      strrep(fill, pad_left),
      txt,
      strrep(fill, pad_right),
      border
    )
  }

  cat(banner_string, "\n")
  # copy banner to allow direct pasting (requires clipr)
  clipr::write_clip(banner_string)
  # avoid [1] when printing if want to manually copy
  invisible(banner_string)
}
#####################################
# function: format p-values for text
# Usage (in-line): `r format_p_vals(ad_test_result$p.value)`
# Usage (console): format_p_vals(ad_test_result$p.value)
#####################################
format_p_vals <- function(p) {
  if (length(p) != 1L || is.na(p)) {
    stop("Error! p must be a single non-missing value")
  }
  if (p > 1) {
    stop("Error! Value greater than 1")
  }
  if (p < 0) {
    stop("Error! Value less than 0")
  }

  if (p >= 0.01) {
    paste0("= ", formatC(p, format = "f", digits = 2))
  } else if (p >= 0.001) {
    paste0("= ", formatC(p, format = "f", digits = 3))
  } else {
    "< 0.001"
  }
}
#####################################
# function: format confidence intervals for tables & text
# Usage (in-line): `r format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])`
# Usage (console): format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])
#####################################
format_interval <- function(lower, upper, digits = 3, small_interval = 5L) {
  paste0(
    "[",
    formatC(
      lower,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    ", ",
    formatC(
      upper,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    "]"
  )
}
#####################################
# function: format truncated ellipses
# Usage (in-line): `r tbi`
# Usage (console): tbi
#####################################
fmt_trunc_ellip <- function(
  x,
  digits = 7,
  ellip = "...",
  tol = 1e-12,
  trim_zeros_if_exact = TRUE
) {
  out <- rep(NA_character_, length(x))
  ok <- is.finite(x)
  scale <- 10^digits
  xt <- trunc(x[ok] * scale) / scale
  s <- formatC(xt, format = "f", digits = digits)
  add <- abs(x[ok] - xt) > tol * pmax(1, abs(x[ok]))
  if (trim_zeros_if_exact) {
    s_trim <- sub("0+$", "", s) # drop trailing zeros
    s_trim <- sub("\\.$", "", s_trim) # drop trailing decimal point
  } else {
    s_trim <- s
  }
  out[ok] <- ifelse(add, paste0(s, ellip), s_trim)
  out[is.na(x)] <- NA_character_
  out
}
#####################################
# function: function to format 6dp numbers with a space after the 3rd decimal
# digit
# Usage (in-line): TBI
# Usage (console): TBI
#####################################
format_spaced_6dp <- function(x) {
  # Format to 6 decimal places
  s <- sprintf("%.6f", x)
  # Insert space between the first 3 and last 3 decimal digits
  sub("(\\.[0-9]{3})([0-9]{3})$", "\\1 \\2", s)
}
```

# Coursework 1 {-}

---

# Shrinkage Estimation for Portfolio Construction

note: want separate R states / instances to allow for no mixing/matching of
variables

## Sub section 2 (want B1 or A1)

### Sub section 3

```{=latex}
\clearpage
\newpage
```

# Coursework 2 {-}

---

# Shrinkage Generalised Linear Models for Insurance Claim Severity Estimation

```{=latex}
% Set up with custom shortened header to avoid overlap
\sectionmark{Shrinkage GLMs for Insurance Claim Severity Estimation}
```

## First part

```{=latex}
\newpage
```

<!-- # Conclusions

TBI -->

<!-- ```{=html} -->
<!-- # References -->
<!-- <div id="refs"></div> -->
<!-- ``` -->

<!-- ```{=latex} -->
<!-- \bibliography{references.bib} -->
<!-- ``` -->

```{=latex}
\newpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
```

# (APPENDIX) Appendices {-}

```{=latex}
\titleformat{\section}[block]{\LARGE\bfseries}{\thesection}{1em}{}
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\thesection-\Alph{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
\setcounter{section}{0}
```

# Supplemental methodology {#apd-supplementary-methods}

TBI

# Reproducibility, accessibility & AI declaration

<!-- BE note: taken out as we are submitting R file separately -->
<!-- Use: `knitr::purl("cs1-group07.rmd", documentation = 0)` to generate r file -->
<!-- Use: `tools::showNonASCIIfile("cs1-group07.rmd")` to debug -->
<!-- ## R Code Documentation -->
<!-- ```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}``` -->

## Reproducibility & accessibility

An accessible HTML version of this report is available via a public GitHub page:

- !Link TBI

This report was created in R Markdown. The source code is open source and is
available via:

- !Link TBI

Changes made to this document were tracked using Git and are also available via
the same repository (!link TBI). Please note that this commit history reflects
updates made to the project repository and is not a comprehensive timeline of
all contributions by group members.

## Generative AI Declaration

Generative AI (GenAI) tools were used throughout this project to assist with
project planning, methodological validation, and debugging. In practice, these
tools functioned as an additional technical team member, providing immediate
feedback, troubleshooting code, and challenging statistical assumptions.

### Tools Used

! To be updated... Google Gemini 3 Pro (Deep Research), ChatGPT 5.2

<!-- ### Prompts used -->

<!-- The GenAI tools were provided with the 'coursework_CM2_250120.pdf' file and -->
<!-- extracts from our `R` code. -->

<!-- A non-exhaustive list of prompts used is shown below. We opted to include at -->
<!-- least one example of each _type_ of prompt used, rather than a list of all -->
<!-- prompts. For example, if we entered the prompts 'How can we put two ggplots side -->
<!-- by side?' and 'How to remove grey background on a ggplot', we would include only -->
<!-- the first of these in the list below, as both relate to plotting issues (& are -->
<!-- similar to 'querying' with a traditional search engine). -->

<!-- - "...the top of both graphs in r are slightly different. (Pasting R code.) Can -->
<!--   this be fixed for the plot area?" -->
<!-- - "Help to identify the key points here" -->
<!-- - "How can this error be tracked down in an rmd file? ! Text line contains an -->
<!--   invalid character. l.1 ^^@^^@..." -->

<!-- ### Use of GenAI outputs & subsequent revisions -->

<!-- The GenAI tools served as a "consultant" member of the group. Notably, the AI -->
<!-- output: -->

<!-- - Helped debug plotting issues. -->
<!-- - Commented on the structure of R code snippets and the report structure. -->

<!-- ### Changes made: -->

<!-- All AI suggestions were critically reviewed and debated by the group before -->
<!-- implementation. For example: -->

<!-- - While the AI suggested code for complex layouts, the specific aesthetic -->
<!--   choices and annotations were manually adjusted to ensure clarity. -->
<!-- - Where the AI proposed specific interpretations, these were cross-referenced -->
<!--   with academic literature and statistics textbooks. -->
<!-- - The final R code for implementation was written and debugged by the group with -->
<!--   AI snippets being treated as drafts rather than final solutions. -->

<!-- While GenAI was treated as a collaborator, the final analysis, statistical -->
<!-- interpretation, and the written content of this report were generated and -->
<!-- verified by the (human) group members. -->

<!-- [^utility-theory]: -->
<!--     Described in more detail in the appendix (\S\@ref(apd-utility-theory)). -->

<!-- [^utility-skew]: -->
<!--     This is demonstrated in the appendix (\S\@ref(apd-utility-theory) and Table -->
<!--     \@ref(tab:exact-utility-table)) where we show that a constant relative risk -->
<!--     aversion (CRRA) utility function with $\gamma > 0$ mathematically confirms -->
<!--     that for risk-averse investors asset 1 results in a lower expected utility -->
<!--     that asset 2. -->
