---
title: "Shrinkage Estimation in Finance and Insurance: Portfolio Construction and Claim Severity Modelling"
subtitle: "SMM069 (2025-26) Group Coursework - Group 12"
author:
  - "Gaurav Agrawal"
  - "Benjamin Evans"
  - "Samyak Jain"
  - "Basmah Khan"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    self_contained: true
    css: style.css
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: paper
    highlight: tango
    df_print: paged
  # runtime: shiny
  bookdown::pdf_document2:
    includes:
      in_header: preamble.tex
    pandoc_args:
      - "--lua-filter=answer.lua"
    number_sections: true
    toc: true
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
# fontsize: 10pt
bibliography: references.bib
---

```{=latex}
\newpage
```

```{=html}
<h1>Information</h1>

This R Markdown document was created as part of a group assignment for SMM069 at Bayes Business School, City St George's, University of London in Term 3 2025-26.

```

```{r asthetic-header, include=FALSE, echo=FALSE}
# ============================================================================ #
# Key Information ====
# SMM069 Advanced Predictive Analytics
# Group Coursework 2025-26
# Group:        Group 12
# Authors (in alphabetical order):
#   - "Gaurav Agrawal"
#   - "Benjamin Evans"
#   - "Samyak Jain"
#   - "Basmah Khan"
# Professor:    Prof Vali Asimit
# Institution:  Bayes Business School - City St George's, University of London
# Date:         09 Jul 2026
#
# Description:  Term 3 group project for SMM069 Advanced Predictive Analtyics
#               (100% of coursework grade - 100% of module grade). The R code
#               below has been exported directly from an R Markdown (.rmd) file.
#               Hence the knitr settings.
#
# Dependencies (!TBI):
#   - tidyverse: loads dplyr, tidyr, ggplot2, purrr (& others)
#   - scales: to express portfolio weights as percentage
#   - patchwork: for combining multiple plots into one figure
#
# Acknowledgements
#   - TBI
# ============================================================================ #
```

```{r setup-knitr, include=FALSE, echo=FALSE}
# Setup & configuration ====
#----------------------- Initial setup (knitr settings) -----------------------#
dir.create("fig", showWarnings = FALSE)

# Defaults common to all outputs
knitr::opts_chunk$set(
  echo     = TRUE,
  message  = FALSE,
  warning  = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.path  = "fig/",
  dpi       = 300
)

# Output-specific settings
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "pdf",
    fig.pos = "!ht",
    out.extra = ""
  )
} else {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "svglite"  # or "png"
  )
}
```

```{r setup-qol, include=FALSE, echo=FALSE}
#----------------------------- Clean environment ------------------------------#
rm(list = ls()) # Remove all objects
graphics.off() # Close all graphical devices
cat("\014") # Clean console
```

```{r load-dependencies, include=FALSE, echo=FALSE}
#------------------- Load dependencies / external libraries -------------------#
# --- Data Manipulation & Plotting ---
library(tidyverse) # loads dplyr, tidyr, ggplot2, purrr (& others)
library(patchwork) # for combining multiple plots into one figure

# --- Reporting & Tables ---
library(kableExtra) # for general tables
library(DT) # for HTML interactive data tables
library(plotly) # for interactive html plots
library(pander) # for writing to pandoc

library(plotly) # interactive html plots

# tmp ====
library(readxl) # reading xlsx GICS sectors
library(PerformanceAnalytics) 
library(xts) 
library(reshape2) 

# Estimator functions ====
source("All_eight_shrinkage_estimators.R")  # St_MSh_, D_MSh_, O_LSh_, T_LSh_, Wang_, BOP_, Jorion_estimator
source("cov1Para.R")     

# Setting random seed ====
set.seed(12) # seed 12 for group 12 ( ˶ˆᗜˆ˵ )
```

```{r html-app, echo=FALSE, results='asis', eval=knitr::is_html_output(), purl=FALSE}
#------------------------------ HTML link to app ------------------------------#
library(htmltools)

# div(style = "background-color: #f8f9fa; padding: 20px; border: 1px solid #e9ecef; border-radius: 5px; text-align: center; margin-bottom: 30px;",
#   h3("Bonus Interactive Dashboard Available"),
#   p("This static report is accompanied by a live R Shiny dashboard allowing one to use a selection of 16 US stocks in different 5 portfolio combinations."),
#   a(href = "https://3enji.shinyapps.io/smm068-a01-g08-dashboard/",
#     target = "_blank",
#     class = "btn btn-primary",
#     style = "background-color: #007bff; color: white; padding: 10px 20px; text-decoration: none; border-radius: 5px; font-weight: bold;",
#     "Launch Interactive Dashboard")
# )
```

```{r custom-functions, include=FALSE, echo=FALSE}
#---------------------------- Custom QOL functions ----------------------------#
#####################################
# function: banner comments (used to to section up code)
# Usage: banner_comment("Element 1: data cleaning") -> then ctrl + v (or cmd+v)
#####################################
banner_comment <- function(text, width = 80, border = "#", fill = "-") {
  txt <- paste0(" ", text, " ")
  inner_width <- width - 2 * nchar(border)
  banner_string <- ""

  if (inner_width <= nchar(txt)) {
    banner_string <- paste0(border, txt, border)
  } else {
    pad_total <- inner_width - nchar(txt)
    pad_left <- pad_total %/% 2
    pad_right <- pad_total - pad_left

    banner_string <- paste0(
      border,
      strrep(fill, pad_left),
      txt,
      strrep(fill, pad_right),
      border
    )
  }

  cat(banner_string, "\n")
  # copy banner to allow direct pasting (requires clipr)
  clipr::write_clip(banner_string)
  # avoid [1] when printing if want to manually copy
  invisible(banner_string)
}
#####################################
# function: format p-values for text
# Usage (in-line): `r format_p_vals(ad_test_result$p.value)`
# Usage (console): format_p_vals(ad_test_result$p.value)
#####################################
format_p_vals <- function(p) {
  if (length(p) != 1L || is.na(p)) {
    stop("Error! p must be a single non-missing value")
  }
  if (p > 1) {
    stop("Error! Value greater than 1")
  }
  if (p < 0) {
    stop("Error! Value less than 0")
  }

  if (p >= 0.01) {
    paste0("= ", formatC(p, format = "f", digits = 2))
  } else if (p >= 0.001) {
    paste0("= ", formatC(p, format = "f", digits = 3))
  } else {
    "< 0.001"
  }
}
#####################################
# function: format confidence intervals for tables & text
# Usage (in-line): `r format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])`
# Usage (console): format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])
#####################################
format_interval <- function(lower, upper, digits = 3, small_interval = 5L) {
  paste0(
    "[",
    formatC(
      lower,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    ", ",
    formatC(
      upper,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    "]"
  )
}
#####################################
# function: format truncated ellipses
# Usage (in-line): `r tbi`
# Usage (console): tbi
#####################################
fmt_trunc_ellip <- function(
  x,
  digits = 7,
  ellip = "...",
  tol = 1e-12,
  trim_zeros_if_exact = TRUE
) {
  out <- rep(NA_character_, length(x))
  ok <- is.finite(x)
  scale <- 10^digits
  xt <- trunc(x[ok] * scale) / scale
  s <- formatC(xt, format = "f", digits = digits)
  add <- abs(x[ok] - xt) > tol * pmax(1, abs(x[ok]))
  if (trim_zeros_if_exact) {
    s_trim <- sub("0+$", "", s) # drop trailing zeros
    s_trim <- sub("\\.$", "", s_trim) # drop trailing decimal point
  } else {
    s_trim <- s
  }
  out[ok] <- ifelse(add, paste0(s, ellip), s_trim)
  out[is.na(x)] <- NA_character_
  out
}
#####################################
# function: function to format 6dp numbers with a space after the 3rd decimal
# digit
# Usage (in-line): TBI
# Usage (console): TBI
#####################################
format_spaced_6dp <- function(x) {
  # Format to 6 decimal places
  s <- sprintf("%.6f", x)
  # Insert space between the first 3 and last 3 decimal digits
  sub("(\\.[0-9]{3})([0-9]{3})$", "\\1 \\2", s)
}
```

# Coursework 1 {-}

---

# Shrinkage estimation for portfolio construction

Aims: Build and evaluate both Mean-Variance portfolios ($\gamma$ = 1) and
Tangency portfolios using sample or shrinkage estimators for the mean and
covariance matrix under a rolling out-of-sample (OOS) setting across the entire
dataset.

## Data preprocessing

We (i) convert the risk-free rate from percentage to decimal, and (ii) compute
weekly excess returns for the 100 companies assigned to Group 12.

```{r a1-load, include=TRUE, echo=FALSE}
# read in data (stored in data folder)
ret_raw <- read.csv(
  "data/DA441_weekly_returns.csv",
  stringsAsFactors = FALSE,
  check.names = FALSE
)
rf_raw <- read.csv("data/risk_free_rate_weekly.csv", stringsAsFactors = FALSE)

# format dates (format is 20090807 -> using %Y%m%d)
ret_raw$Date <- as.Date(as.character(ret_raw$Date), format = "%Y%m%d")
rf_raw$Date <- as.Date(as.character(rf_raw$Date), format = "%Y%m%d")

# (i) convert risk-free rate from percentage -> decimal.
rf_cpy <- rf_raw
rf_cpy$RF <- rf_raw$RF / 100

# align rf to the return dates by date
dat <- merge(ret_raw, rf_cpy[, c("Date", "RF")], by = "Date")
dates <- dat$Date
# original tickers, kept for GICS
ticker_cols <- setdiff(colnames(ret_raw), "Date") 
returns_all <- as.matrix(dat[, ticker_cols])
rf <- dat$RF

# setting seed again
set.seed(12)
selected <- sample(ticker_cols, 100)
returns100 <- returns_all[, selected]

# (ii) Weekly excess returns: r_excess[t,i] = r[t,i] - rf[t].
# sweep() subtracts the rf vector across rows explicitly (clearer than returns100 - rf).
excess_returns <- sweep(returns100, 1, rf, FUN = "-")

cat("Weeks:", nrow(excess_returns), " | Companies:", ncol(excess_returns), "\n")
cat("Date range:", as.character(min(dates)), "to", as.character(max(dates)), "\n")
```

## Selection of estimation models

Using all eight mean-vector estimators (Sample mean, St-MSh, D-MSh, O-LSh,
T-LSh, Wang, BOP, Jorion) and a single covariance estimator: `cov1Para` (check
full name of this from lecture slides).

### Why `cov1Para`? (vs sample covariance)

Each training window has $n = 156$ weeks for $p = 100$ assets, 
$$\rightarrow\frac{n}{p} \approx 1.56$$

<!-- The sample covariance -->
<!-- is technically invertible ($n > p$) but poorly conditioned at this ratio: its -->
<!-- smallest eigenvalues are badly underestimated, the inverse blows them up, and the -->
<!-- resulting Mean-Variance / Tangency weights become extreme and unstable out of -->
<!-- sample. `cov1Para` shrinks the sample covariance toward a scaled identity by an -->
<!-- optimally estimated intensity, guaranteeing a well-conditioned, invertible matrix -->
<!-- and far more stable portfolio weights - exactly the regime this estimator was -->
<!-- designed for. We therefore use it for every portfolio below. -->

```{r a2-estimators, include=FALSE, echo=FALSE}
# Mean-vector estimators 
# Check this* (each takes a returns matrix: rows = weeks, cols = assets).
mean_estimators <- list(
  Sample = function(X) colMeans(X),
  StM    = St_MSh_estimator,
  DM     = D_MSh_estimator,
  OL     = O_LSh_estimator,
  TL     = T_LSh_estimator,
  Wang   = Wang_estimator,
  BOP    = BOP_estimator,
  Jorion = Jorion_estimator
)

# Single covariance estimator for the whole analysis
cov_fun <- cov1Para
```

## Portfolio construction and rolling-window setup

Training = 156 weeks, testing = 4 weeks, step = 4 weeks. 

For each window and each
mean estimator we build a Mean-Variance portfolio (omega = 1) and a Tangency
portfolio; we also include the Equal-Weight benchmark (EWP).

```{r a3-windows, include=TRUE, echo=FALSE}
training_weeks <- 156
testing_weeks <- 4
step_weeks <- 4
omega <- 1

n_total <- nrow(excess_returns)
n_windows <- floor((n_total - training_weeks - testing_weeks) / step_weeks) + 1
cat("Total number of rolling windows:", n_windows, "\n")
```

```{r a3-helpers, include=FALSE, echo=FALSE}
# annualised Sharpe ratio from a weekly excess-return series: 
# (mean/sd) * sqrt(52).
ann_sharpe <- function(r, scale = 52) {
  r <- as.numeric(r)
  s <- sd(r)
  if (is.na(s) || s == 0) {
    return(NA_real_)
  }
  (mean(r) / s) * sqrt(scale)
}

# Evaluate one window: returns a named vector of annualised Sharpe ratios for
# MV_<mean>, TP_<mean> (8 each) and EWP -> 17 strategies.
eval_window <- function(train_X, test_X, cov_fun, omega = 1) {
  Sigma <- cov_fun(train_X)
  Sigma_inv <- solve(Sigma)
  out <- numeric(0)

  for (nm in names(mean_estimators)) {
    mu <- mean_estimators[[nm]](train_X)
    z <- as.vector(Sigma_inv %*% mu) # unnormalised Sigma^{-1} mu

    w_mv <- (1 / omega) * z # Mean-Variance (omega = 1): NOT normalised
    w_tp <- z / sum(z) # Tangency: fully invested, weights sum to 1

    # Portfolio excess-return series over the 4-week test window.
    out[paste0("MV_", nm)] <- ann_sharpe(test_X %*% w_mv)
    out[paste0("TP_", nm)] <- ann_sharpe(test_X %*% w_tp)
  }

  # Equal-Weight benchmark.
  p <- ncol(train_X)
  w_ew <- rep(1 / p, p)
  out["EWP"] <- ann_sharpe(test_X %*% w_ew)

  out
}
```

```{r a3-rolling, include=FALSE, echo=FALSE}
sr_list <- vector("list", n_windows)
test_start_dates <- as.Date(rep(NA, n_windows))

for (k in seq_len(n_windows)) {
  tr_start <- 1 + (k - 1) * step_weeks
  tr_end <- tr_start + training_weeks - 1
  te_start <- tr_end + 1
  te_end <- te_start + testing_weeks - 1

  train_X <- excess_returns[tr_start:tr_end, , drop = FALSE]
  test_X <- excess_returns[te_start:te_end, , drop = FALSE]

  sr_list[[k]] <- eval_window(train_X, test_X, cov_fun = cov_fun, omega = omega)
  test_start_dates[k] <- dates[te_start]
}

sr_mat <- do.call(rbind, sr_list)
sr_df <- data.frame(TestStart = test_start_dates, sr_mat, check.names = FALSE)
```

## Portfolio performance

Annualised Sharpe per window, summarised (mean and SD) over three environments.
Windows are assigned to a sub-period by their test-window start date.

<!-- Alternatives: require the whole test window inside the range, or use te_end. -->

```{r a4-subperiods, include=FALSE, echo=FALSE}
fc_lo <- as.Date("2008-10-06")
fc_hi <- as.Date("2011-08-03") # Financial Crisis
pd_lo <- as.Date("2020-02-26")
pd_hi <- as.Date("2023-12-29") # Pandemic

masks <- list(
  Full = rep(TRUE, nrow(sr_df)),
  FinancialCrisis = sr_df$TestStart >= fc_lo & sr_df$TestStart <= fc_hi,
  Pandemic = sr_df$TestStart >= pd_lo & sr_df$TestStart <= pd_hi
)

strategy_cols <- setdiff(colnames(sr_df), "TestStart")

summarise_period <- function(mask) {
  sub <- sr_df[mask, strategy_cols, drop = FALSE]
  data.frame(
    Mean = sapply(sub, mean, na.rm = TRUE),
    SD = sapply(sub, sd, na.rm = TRUE)
  )
}

summary_full <- summarise_period(masks$Full)
summary_fc <- summarise_period(masks$FinancialCrisis)
summary_pd <- summarise_period(masks$Pandemic)

cat(
  "Windows -> Full:",
  sum(masks$Full),
  "| Financial Crisis:",
  sum(masks$FinancialCrisis),
  "| Pandemic:",
  sum(masks$Pandemic),
  "\n"
)
```

```{r a4-tables, include=TRUE, echo=FALSE}
kable(
  summary_full[order(-summary_full$Mean), ],
  digits = 4,
  caption = "Annualised Sharpe Ratio - Full Period (mean, SD)",
  booktabs = TRUE
) %>%
  kable_styling(
    full_width = FALSE,
    position   = "center"
  ) 

kable(
  summary_fc[order(-summary_fc$Mean), ],
  digits = 4,
  caption = "Annualised Sharpe Ratio - Financial Crisis (mean, SD)",
  booktabs = TRUE
) %>%
  kable_styling(
    full_width = FALSE,
    position   = "center"
  ) 

kable(
  summary_pd[order(-summary_pd$Mean), ],
  digits = 4,
  caption = "Annualised Sharpe Ratio - Pandemic (mean, SD)",
  booktabs = TRUE
) %>%
  kable_styling(
    full_width = FALSE,
    position   = "center"
  ) 

```

```{r a4-gics, include=TRUE, echo=FALSE}
gics <- as.data.frame(read_excel("data/DA441_GICS_Sectors.xlsx"))

# names(gics) # headers
id_col <- "DA441_column_name"
sector_col <- "GICS Sector Name"

sel_gics <- gics[gics[[id_col]] %in% selected, ]
nrow(sel_gics) # check: should be 100 (Y)

sector_tab <- as.data.frame(table(sel_gics[[sector_col]]))
colnames(sector_tab) <- c("Sector", "Count") 
sector_tab <- sector_tab[order(-sector_tab$Count), ]
```

```{r a4-gics-table, include=TRUE, echo=FALSE}
kable(
  sector_tab,
  row.names = FALSE,
  booktabs = TRUE,
  caption = "GICS sector composition of the 100 sampled companies (Group 12)"
) %>%
  kable_styling(full_width = FALSE, position = "center")
```

```{r a4-gics-figures, include=TRUE, echo=FALSE}
# Doughnut chart
okabe_ito <- c("#E69F00","#56B4E9","#009E73","#F0E442",
               "#0072B2","#D55E00","#CC79A7","#FFFFF0",
               "#999999", "#D4AEE3", "#B8E986")

p_donut <- ggplot(sector_tab, aes(x = 2, y = Count, fill = Sector)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  geom_text(
    aes(label = paste0(round(100 * Count / sum(Count), 1), "%")),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  theme_void() +
  scale_fill_manual(values = okabe_ito) +
  labs(
    title = "GICS sector composition (Group 12, 100 companies)",
    fill = "GICS sector"
  )

if (knitr::is_html_output()) {
  # plotly in html
  plot_ly(
    data = sector_tab, 
    labels = ~Sector, 
    values = ~Count, 
    type = 'pie', 
    hole = 0.6, 
    marker = list(colors = okabe_ito),
    textinfo = 'label+percent' 
  ) %>%
    layout(
      title = "GICS sector composition (Group 12, 100 companies)",
      showlegend = FALSE 
    )
} else {
  p_donut
}

p_bar <- ggplot(sector_tab, aes(x = Count, y = reorder(Sector, Count), text = paste0(round(100 * Count / sum(Count), 1), "%"))) +
  geom_col(fill = "#0072B2") +
  labs(
    title = "GICS sector composition (Group 12, 100 companies)",
    x = "Companies",
    y = NULL
  ) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 100, by = 5)) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank()
  )

# Conditional Rendering for Bar
if (knitr::is_html_output()) {
  ggplotly(p_bar, tooltip = c("y", "x", "text"))
} else {
  p_bar
}
```

## Interpretation of results and final report

```{=latex}
\clearpage
\newpage
```

## Rough working

```{r, echo=FALSE, include=FALSE}
u <- 2.3212
d <- 3.113131
```

<!-- prettier-ignore -->
::: {.answer-wrapper}
::: {.answer}
$u=`r formatC(u, format="f", digits=5)`, \quad d=`r formatC(d, format="f", digits=5)`$
:::
:::

### Sub section 3

```{=latex}
\clearpage
\newpage
```

# Coursework 2 {-}

---

# Shrinkage Generalised Linear Models for Insurance Claim Severity Estimation

```{=latex}
% Set up with custom shortened header to avoid overlap
\sectionmark{Shrinkage GLMs for Insurance Claim Severity Estimation}
```

## First part

```{=latex}
\newpage
```

<!-- # Conclusions

TBI -->

<!-- ```{=html} -->
<!-- # References -->
<!-- <div id="refs"></div> -->
<!-- ``` -->

<!-- ```{=latex} -->
<!-- \bibliography{references.bib} -->
<!-- ``` -->

```{=latex}
\newpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
```

# (APPENDIX) Appendices {-}

```{=latex}
\titleformat{\section}[block]{\LARGE\bfseries}{\thesection}{1em}{}
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\thesection-\Alph{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
\setcounter{section}{0}
```

# Supplemental methodology {#apd-supplementary-methods}

TBI

# Reproducibility, accessibility & AI declaration

<!-- BE note: taken out as we are submitting R file separately -->
<!-- Use: `knitr::purl("cs1-group07.rmd", documentation = 0)` to generate r file -->
<!-- Use: `tools::showNonASCIIfile("cs1-group07.rmd")` to debug -->
<!-- ## R Code Documentation -->
<!-- ```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}``` -->

## Reproducibility & accessibility

An accessible HTML version of this report is available via a public GitHub page:

- !Link TBI

This report was created in R Markdown. The source code is open source and is
available via:

- !Link TBI

Changes made to this document were tracked using Git and are also available via
the same repository (!link TBI). Please note that this commit history reflects
updates made to the project repository and is not a comprehensive timeline of
all contributions by group members.

## Generative AI Declaration

Generative AI (GenAI) tools were used throughout this project to assist with
project planning, methodological validation, and debugging. In practice, these
tools functioned as an additional technical team member, providing immediate
feedback, troubleshooting code, and challenging statistical assumptions.

### Tools Used

! To be updated... Google Gemini 3 Pro (Deep Research), ChatGPT 5.2

<!-- ### Prompts used -->

<!-- The GenAI tools were provided with the 'coursework_CM2_250120.pdf' file and -->
<!-- extracts from our `R` code. -->

<!-- A non-exhaustive list of prompts used is shown below. We opted to include at -->
<!-- least one example of each _type_ of prompt used, rather than a list of all -->
<!-- prompts. For example, if we entered the prompts 'How can we put two ggplots side -->
<!-- by side?' and 'How to remove grey background on a ggplot', we would include only -->
<!-- the first of these in the list below, as both relate to plotting issues (& are -->
<!-- similar to 'querying' with a traditional search engine). -->

<!-- - "...the top of both graphs in r are slightly different. (Pasting R code.) Can -->
<!--   this be fixed for the plot area?" -->
<!-- - "Help to identify the key points here" -->
<!-- - "How can this error be tracked down in an rmd file? ! Text line contains an -->
<!--   invalid character. l.1 ^^@^^@..." -->

<!-- ### Use of GenAI outputs & subsequent revisions -->

<!-- The GenAI tools served as a "consultant" member of the group. Notably, the AI -->
<!-- output: -->

<!-- - Helped debug plotting issues. -->
<!-- - Commented on the structure of R code snippets and the report structure. -->

<!-- ### Changes made: -->

<!-- All AI suggestions were critically reviewed and debated by the group before -->
<!-- implementation. For example: -->

<!-- - While the AI suggested code for complex layouts, the specific aesthetic -->
<!--   choices and annotations were manually adjusted to ensure clarity. -->
<!-- - Where the AI proposed specific interpretations, these were cross-referenced -->
<!--   with academic literature and statistics textbooks. -->
<!-- - The final R code for implementation was written and debugged by the group with -->
<!--   AI snippets being treated as drafts rather than final solutions. -->

<!-- While GenAI was treated as a collaborator, the final analysis, statistical -->
<!-- interpretation, and the written content of this report were generated and -->
<!-- verified by the (human) group members. -->

<!-- [^utility-theory]: -->
<!--     Described in more detail in the appendix (\S\@ref(apd-utility-theory)). -->

<!-- [^utility-skew]: -->
<!--     This is demonstrated in the appendix (\S\@ref(apd-utility-theory) and Table -->
<!--     \@ref(tab:exact-utility-table)) where we show that a constant relative risk -->
<!--     aversion (CRRA) utility function with $\gamma > 0$ mathematically confirms -->
<!--     that for risk-averse investors asset 1 results in a lower expected utility -->
<!--     that asset 2. -->
