---
title:
  "Shrinkage Estimation in Finance and Insurance: Portfolio Construction and
  Claim Severity Modelling"
subtitle: "SMM069 (2025-26) Group Coursework - Group 12"
author:
  - "Gaurav Agrawal"
  - "Benjamin Evans"
  - "Samyak Jain"
  - "Basmah Khan"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    self_contained: true
    css: style.css
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
    theme: paper
    highlight: tango
    df_print: paged
    # runtime: shiny
    pandoc_args:
      - "--lua-filter=style_code.lua"
  bookdown::pdf_document2:
    includes:
      in_header: preamble.tex
    pandoc_args:
      - "--lua-filter=answer.lua"
      - "--lua-filter=style_code.lua"
    number_sections: true
    toc: true
    latex_engine: xelatex
    citation_package: natbib
    keep_tex: true
# fontsize: 10pt
bibliography: references.bib
---

```{=latex}
\newpage
```

```{=html}
<h1>Information</h1>

This R Markdown document was created as part of a group assignment for SMM069 at Bayes Business School, City St George's, University of London in Term 3 2025-26.

```

```{r asthetic-header, include=FALSE, echo=FALSE}
# ============================================================================ #
# Key Information ====
# SMM069 Advanced Predictive Analytics
# Group Coursework 2025-26
# Group:        Group 12
# Authors (in alphabetical order):
#   - "Gaurav Agrawal"
#   - "Benjamin Evans"
#   - "Samyak Jain"
#   - "Basmah Khan"
# Professor:    Prof Vali Asimit & Ms Ziwei Chen
# Institution:  Bayes Business School - City St George's, University of London
# Date:         09 Jul 2026
#
# Description:  Term 3 group project for SMM069 Advanced Predictive Analtyics
#               (100% of coursework grade - 100% of module grade).
#
# Dependencies:
#   - TBI
#
#   - tidyverse: loads dplyr, tidyr, ggplot2, purrr (& others)
#   - scales: to express portfolio weights as percentage
#   - patchwork: for combining multiple plots into one figure
#
# Acknowledgements
#   - TBI
#
# Note to Vali and Ziwei:
#   We originally planned to submit just an R file exported from this RMD,
#   hence this 'asthetic header'. We left it in as a legacy artefact after
#   deciding to submit our code as this single RMD file.
# ============================================================================ #
```

```{r setup-knitr, include=FALSE, echo=FALSE}
# Setup & configuration ====
#----------------------- Initial setup (knitr settings) -----------------------#
dir.create("fig", showWarnings = FALSE)

# Defaults common to all outputs
knitr::opts_chunk$set(
  echo     = FALSE,
  message  = FALSE,
  warning  = FALSE,
  fig.align = "center",
  out.width = "100%",
  fig.path  = "fig/",
  dpi       = 300
)

# Output-specific settings
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "pdf",
    fig.pos = "!ht",
    out.extra = ""
  )
} else {
  knitr::opts_chunk$set(
    fig.width = 6,
    fig.height = 4,
    dev = "svglite"  # or "png"
  )
}
```

```{r setup-qol, include=FALSE, echo=FALSE}
#----------------------------- Clean environment ------------------------------#
rm(list = ls()) # Remove all objects
graphics.off() # Close all graphical devices
cat("\014") # Clean console
```

```{r load-dependencies, include=FALSE, echo=FALSE}
#------------------- Load dependencies / external libraries -------------------#
# --- Data Manipulation & Plotting ---
library(tidyverse) # loads dplyr, tidyr, ggplot2, purrr (& others)
library(patchwork) # for combining multiple plots into one figure

# --- Reporting & Tables ---
library(kableExtra) # for general tables
library(DT)         # for HTML interactive data tables
library(plotly)     # for interactive html plots
library(pander)     # for writing to pandoc

# --- Statistical Modelling & Mathematics ---
library(MASS)      # Support functions and datasets for applied statistics
library(glm2)      # Fitting generalised linear models
library(glmnet)    # Lasso and elastic-net
library(splines)   # Regression spline functions

# --- Course-Specific / Custom Packages ---
library(savvySh)   # Custom coursework package
library(savvyPR)   # Custom coursework package
library(savvyGLM)  # Custom coursework package

# tmp ====
library(readxl) # reading xlsx GICS sectors
library(PerformanceAnalytics)
library(xts)
library(reshape2)

# Estimator functions ====
# St_MSh_, D_MSh_, O_LSh_, T_LSh_, Wang_, BOP_, Jorion_estimator
source("All_eight_shrinkage_estimators.R")
source("cov1Para.R")

# Global env options ====
## Setting random seed ----
group_num <- 12 # seed 12 for group 12 ( ˶ˆᗜˆ˵ )
set.seed(group_num)
## disable scientific notation ----
options(scipen = 999) 
```

```{r custom-functions, include=FALSE, echo=FALSE}
#---------------------------- Custom QOL functions ----------------------------#
# BE Note: most of these are legacy functions from previous projects - just
# keep them here to have all in one contained file.
#####################################
# function: format p-values for text
# Usage (in-line): `r format_p_vals(ad_test_result$p.value)`
# Usage (console): format_p_vals(ad_test_result$p.value)
#####################################
format_p_vals <- function(p) {
  if (length(p) != 1L || is.na(p)) {
    stop("Error! p must be a single non-missing value")
  }
  if (p > 1) {
    stop("Error! Value greater than 1")
  }
  if (p < 0) {
    stop("Error! Value less than 0")
  }

  if (p >= 0.01) {
    paste0("= ", formatC(p, format = "f", digits = 2))
  } else if (p >= 0.001) {
    paste0("= ", formatC(p, format = "f", digits = 3))
  } else {
    "< 0.001"
  }
}
#####################################
# function: format confidence intervals for tables & text
# Usage (in-line): `r format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])`
# Usage (console): format_interval(el2_ci_normal_95[1], el2_ci_normal_95[2])
#####################################
format_interval <- function(lower, upper, digits = 3, small_interval = 5L) {
  paste0(
    "[",
    formatC(
      lower,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    ", ",
    formatC(
      upper,
      format = "f",
      digits = digits,
      small.interval = small_interval,
      small.mark = " "
    ),
    "]"
  )
}
#####################################
# function: format truncated ellipses
# Usage (in-line): `r tbi`
# Usage (console): tbi
#####################################
fmt_trunc_ellip <- function(
  x,
  digits = 7,
  ellip = "...",
  tol = 1e-12,
  trim_zeros_if_exact = TRUE
) {
  out <- rep(NA_character_, length(x))
  ok <- is.finite(x)
  scale <- 10^digits
  xt <- trunc(x[ok] * scale) / scale
  s <- formatC(xt, format = "f", digits = digits)
  add <- abs(x[ok] - xt) > tol * pmax(1, abs(x[ok]))
  if (trim_zeros_if_exact) {
    s_trim <- sub("0+$", "", s) # drop trailing zeros
    s_trim <- sub("\\.$", "", s_trim) # drop trailing decimal point
  } else {
    s_trim <- s
  }
  out[ok] <- ifelse(add, paste0(s, ellip), s_trim)
  out[is.na(x)] <- NA_character_
  out
}
#####################################
# function: function to format 6dp numbers with a space after the 3rd decimal
# digit
# Usage (in-line): TBI
# Usage (console): TBI
#####################################
format_spaced_6dp <- function(x) {
  # Format to 6 decimal places
  s <- sprintf("%.6f", x)
  # Insert space between the first 3 and last 3 decimal digits
  sub("(\\.[0-9]{3})([0-9]{3})$", "\\1 \\2", s)
}
#####################################
# function: compact inline number formatter
# Usage (in-line): `r fmt(2.32320)`
fmt <- function(x, d = 2) {
  formatC(x, format = "f", digits = d, big.mark = ",")
}
```

# Coursework 1 {-}

# Shrinkage estimation for portfolio construction

```{r a-data-load, include=FALSE}
no_companies <- 100
# read in data (stored in data folder)
ret_raw <- read.csv(                    
  "data/DA441_weekly_returns.csv",
  stringsAsFactors = FALSE, 
  check.names = FALSE
)
rf_raw <- read.csv("data/risk_free_rate_weekly.csv", stringsAsFactors = FALSE)
gics <- as.data.frame(read_excel("data/DA441_GICS_Sectors.xlsx"))
# format dates
ret_raw$Date <- as.Date(as.character(ret_raw$Date), format = "%Y%m%d") 
rf_raw$Date  <- as.Date(as.character(rf_raw$Date),  format = "%Y%m%d")
```

```{r a-data-prep, include=FALSE}
## (i) convert risk-free rate from percentage -> decimal  ----
rf_dec    <- rf_raw # make quick copy
rf_dec$RF <- rf_raw$RF / 100
# align risk free rate to the return dates by date
dat <- merge(ret_raw, rf_dec[, c("Date", "RF")], by = "Date")
dates <- dat$Date
# original tickers, kept for GICS
ticker_cols <- setdiff(colnames(ret_raw), "Date")
returns_all <- as.matrix(dat[, ticker_cols])
rf <- dat$RF
# gets our specific tickers
set.seed(group_num)
selected <- sample(ticker_cols, no_companies)
returns100 <- returns_all[, selected]

## (ii) Weekly excess returns: r_excess[t,i] = r[t,i] - rf[t] ----
# sweep() subtracts the rf vector across rows (clearer than returns100 - rf)
excess_returns <- sweep(returns100, 1, rf, FUN = "-")
```

```{r a-params-estimators, include=FALSE}
# rolling-window parameters
training_weeks  <- 156
testing_weeks   <- 4
step_weeks      <- 4
omega           <- 1      # risk aversion gamma = 1

# BE note: n_assets should always equal no_companies as data comes pre-cleaned
n_total   <- nrow(excess_returns)
n_windows <- floor((n_total - training_weeks - testing_weeks) / step_weeks) + 1
n_weeks   <- n_total
n_assets  <- ncol(excess_returns)
np_ratio  <- training_weeks / n_assets
cat("Total number of rolling windows:", n_windows, "\n")

# Mean-vector estimators
# Check this* (each takes a returns matrix: rows = weeks, cols = assets).
mean_estimators <- list(
  Sample = function(X) colMeans(X),
  StMSh = St_MSh_estimator,
  DMSh = D_MSh_estimator,
  OLSh = O_LSh_estimator,
  TLSh = T_LSh_estimator,
  Wang = Wang_estimator,
  BOP = BOP_estimator,
  Jorion = Jorion_estimator
)
# Mirroring paper ordering from Asimit et al. 2026, Table 3
# proposed estimators -> sample -> benchmarks
est_order <- c(
  "StMSh", "DMSh", "OLSh", "TLSh", "Sample", "Wang", "BOP", "Jorion"
)
stopifnot(setequal(est_order, names(mean_estimators)))
# Single covariance estimator for the whole analysis
cov_fun <- cov1Para

# annualised Sharpe ratio from a weekly excess-return series
# > (mean/sd) * sqrt(52).
# > catches if assets in 4-week window have zero volatility
# > would get `Inf` or `NaN` -> this changes these values to NA()
ann_sharpe <- function(r, scale = 52) {
  r <- as.numeric(r)
  s <- sd(r)
  if (is.na(s) || s == 0) {
    return(NA_real_)
  }
  (mean(r) / s) * sqrt(scale)
}
```

```{r a-cov-diagnostic, include=TRUE}
# conditioning diagnostic
# cov1Para vs sample covariance
# mean estimator and porfolio construction held constant & just sigma changes

run_diag <- function(X, cov_estimator, mean_estimators,
                     training_weeks, testing_weeks, step_weeks) {
  n_windows <- floor((nrow(X) - training_weeks - testing_weeks) / step_weeks) + 1
  conds <- numeric(n_windows)
  Wl <- vector("list", n_windows)
  Rl <- vector("list", n_windows)
  for (k in seq_len(n_windows)) {
    tr_start <- 1 + (k - 1) * step_weeks
    tr_end <- tr_start + training_weeks - 1
    te_start <- tr_end + 1
    te_end <- te_start + testing_weeks - 1
    train_X <- X[tr_start:tr_end, , drop = FALSE]
    test_X <- X[te_start:te_end, , drop = FALSE]

    Sigma <- cov_estimator(train_X)
    ev <- eigen(Sigma, symmetric = TRUE, only.values = TRUE)$values
    conds[k] <- max(ev) / min(ev)
    Sigma_inv <- solve(Sigma)

    W <- sapply(mean_estimators, function(f) {
      z <- as.vector(Sigma_inv %*% f(train_X))
      z / sum(z) # tangency weights
    })
    colnames(W) <- names(mean_estimators)
    Wl[[k]] <- W
    Rl[[k]] <- test_X %*% W
  }
  gross <- rowMeans(sapply(Wl, function(W) colSums(abs(W))))
  turn <- rowMeans(sapply(2:n_windows, function(k) colSums(abs(Wl[[k]] - Wl[[k - 1]]))))
  oos <- apply(do.call(rbind, Rl), 2, ann_sharpe)
  list(
    cond_mean = mean(conds), cond_med = median(conds),
    tab = data.frame(
      GrossLev = gross, Turnover = turn, OOS_Sharpe = oos,
      row.names = names(mean_estimators)
    )
  )
}
diag_shrink <- run_diag(
  excess_returns, cov1Para, mean_estimators,
  training_weeks, testing_weeks, step_weeks
)
diag_sample <- run_diag(
  excess_returns, function(X) cov(X), mean_estimators,
  training_weeks, testing_weeks, step_weeks
)

cond_mean_shrink <- diag_shrink$cond_mean
cond_mean_sample <- diag_sample$cond_mean
cond_med_shrink <- diag_shrink$cond_med
cond_med_sample <- diag_sample$cond_med
cond_ratio_mean <- cond_mean_sample / cond_mean_shrink

cmp <- merge(diag_shrink$tab, diag_sample$tab,
  by = "row.names",
  suffixes = c("_cov1Para", "_sample")
)
rownames(cmp) <- cmp$Row.names
cmp$Row.names <- NULL
# cond_tab <- cmp[names(mean_estimators), ]
cond_tab <- cmp[est_order, ]
```

```{r a-rolling, include=FALSE}
# build mean variance and tangency portfolio weights per estimator (& EWP)
# > Score each by annualised Sharpe ratio over 4 week test window
# > Then roll forward across sample (described in labs)
eval_window <- function(train_X, test_X, cov_fun, mean_estimators, omega = 1) {
  Sigma_inv <- solve(cov_fun(train_X))
  out <- numeric(0)
  for (nm in names(mean_estimators)) {
    mu <- mean_estimators[[nm]](train_X)
    # Sigma^{-1} mu
    z <- as.vector(Sigma_inv %*% mu)
    # mean-variance (gamma = 1) <-> un-normalised
    w_mv <- (1 / omega) * z
    # tangency -> weights sum to 1
    w_tp <- z / sum(z)
    out[paste0("MV_", nm)] <- ann_sharpe(test_X %*% w_mv)
    out[paste0("TP_", nm)] <- ann_sharpe(test_X %*% w_tp)
  }
  w_ew <- rep(1 / ncol(train_X), ncol(train_X))
  # equal-weight benchmark
  out["EWP"] <- ann_sharpe(test_X %*% w_ew)
  out
}

sr_list <- vector("list", n_windows)
test_start_dates <- as.Date(rep(NA, n_windows))
for (k in seq_len(n_windows)) {
  tr_start <- 1 + (k - 1) * step_weeks
  tr_end <- tr_start + training_weeks - 1
  te_start <- tr_end + 1
  te_end <- te_start + testing_weeks - 1
  train_X <- excess_returns[tr_start:tr_end, , drop = FALSE]
  test_X <- excess_returns[te_start:te_end, , drop = FALSE]
  sr_list[[k]] <- eval_window(train_X, test_X,
    cov_fun         = cov_fun,
    mean_estimators = mean_estimators,
    omega           = omega
  )
  test_start_dates[k] <- dates[te_start]
}
sr_mat <- do.call(rbind, sr_list)
sr_df <- data.frame(TestStart = test_start_dates, sr_mat, check.names = FALSE)
```

```{r a-summaries-metrics, include=FALSE}
# Sub-period masks (assigned by test-window start date)
fc_lo <- as.Date("2008-10-06")
fc_hi <- as.Date("2011-08-03") # Financial Crisis
pd_lo <- as.Date("2020-02-26")
pd_hi <- as.Date("2023-12-29") # Pandemic

masks <- list(
  Full            = rep(TRUE, nrow(sr_df)),
  FinancialCrisis = sr_df$TestStart >= fc_lo & sr_df$TestStart <= fc_hi,
  Pandemic        = sr_df$TestStart >= pd_lo & sr_df$TestStart <= pd_hi
)
n_full <- sum(masks$Full)
n_fc <- sum(masks$FinancialCrisis)
n_pd <- sum(masks$Pandemic)

strategy_cols <- setdiff(colnames(sr_df), "TestStart")
summarise_period <- function(df, mask, cols) {
  sub <- df[mask, cols, drop = FALSE]
  data.frame(
    Mean = sapply(sub, mean, na.rm = TRUE),
    SD = sapply(sub, sd, na.rm = TRUE)
  )
}
summary_full <- summarise_period(sr_df, masks$Full, strategy_cols)
summary_fc <- summarise_period(sr_df, masks$FinancialCrisis, strategy_cols)
summary_pd <- summarise_period(sr_df, masks$Pandemic, strategy_cols)

cat(
  "Windows -> Full:",
  sum(masks$Full),
  "| Financial Crisis:",
  sum(masks$FinancialCrisis),
  "| Pandemic:",
  sum(masks$Pandemic),
  "\n"
)

summary_full$Strategy <- rownames(summary_full)
summary_fc$Strategy <- rownames(summary_fc)
summary_pd$Strategy <- rownames(summary_pd)
combined_summary <- summary_full |>
  dplyr::left_join(summary_fc, by = "Strategy", suffix = c("_full", "_fc")) |>
  dplyr::left_join(summary_pd, by = "Strategy") |>
  dplyr::rename(Mean_pd = Mean, SD_pd = SD) |>
  dplyr::relocate("Strategy")

row_order_hl <- c("EWP", paste0("MV_", est_order), paste0("TP_", est_order))
combined_summary <- combined_summary[
  match(row_order_hl, combined_summary$Strategy),
]

# per window mean-variance vs tagency portfolio equality (up to sign)
# & St-MSh == Sample check
mv <- as.matrix(sr_df[, grep("^MV_", names(sr_df))])
tp <- as.matrix(sr_df[, grep("^TP_", names(sr_df))])
mvtp_maxdiff <- max(abs(abs(mv) - abs(tp)), na.rm = TRUE)
stm_check <- isTRUE(all.equal(sr_df$MV_Sample, sr_df$MV_StMSh))

# ========================================================================== #
# scale sensitive out of sample metrics
# -> created as companion to the headline table
oos_metrics <- function(X, mean_estimators, cov_fun,
                        training_weeks, testing_weeks, step_weeks, omega = 1) {
  n_windows <- floor((nrow(X) - training_weeks - testing_weeks) / step_weeks) + 1
  est_names <- names(mean_estimators)
  strat <- c(paste0("MV_", est_names), paste0("TP_", est_names), "EWP")
  p <- ncol(X)

  # [1] Assets (p), [2] Strategy Name, [3] Rolling Window Step
  W_arr <- array(NA_real_,
    dim = c(p, length(strat), n_windows),
    dimnames = list(NULL, strat, NULL)
  )
  R_arr <- array(NA_real_,
    dim = c(testing_weeks, length(strat), n_windows),
    dimnames = list(NULL, strat, NULL)
  )
  sumz <- matrix(NA_real_, n_windows, length(est_names),
    dimnames = list(NULL, est_names)
  )
  for (k in seq_len(n_windows)) {
    tr_start <- 1 + (k - 1) * step_weeks
    tr_end <- tr_start + training_weeks - 1
    te_start <- tr_end + 1
    te_end <- te_start + testing_weeks - 1
    train_X <- X[tr_start:tr_end, , drop = FALSE]
    test_X <- X[te_start:te_end, , drop = FALSE]
    Sigma_inv <- solve(cov_fun(train_X))

    # Z = \Sigma^{-1} \mu
    Z <- sapply(mean_estimators, function(f) as.vector(Sigma_inv %*% f(train_X))) # p x 8
    sumz[k, ] <- colSums(Z)

    # W_mv = Z / \omega  |  W_tp = Z / (\mathbf{1}^T Z)
    W <- cbind(
      Z / omega, # MV (gamma = 1), unnormalised
      sweep(Z, 2, colSums(Z), "/"), # TP, weights sum to 1
      rep(1 / p, p)
    ) # EWP
    colnames(W) <- strat
    W_arr[, , k] <- W
    R_arr[, , k] <- test_X %*% W
  }
  list(W = W_arr, R = R_arr, sumz = sumz, strat = strat)
}
summarise_oos <- function(res, mask = rep(TRUE, dim(res$R)[3]), omega = 1, w_stat = mean) {
  W <- res$W[, , mask, drop = FALSE]
  R <- apply(res$R[, , mask, drop = FALSE], 2, c) # stitched weekly series
  ann_ret <- 52 * colMeans(R)
  ann_vol <- sqrt(52) * apply(R, 2, sd)
  sharpe <- apply(R, 2, ann_sharpe)
  ce <- ann_ret - (omega / 2) * ann_vol^2 # realised certainty-equivalent
  gross <- apply(apply(abs(W), c(2, 3), sum), 1, w_stat)
  maxw <- apply(apply(abs(W), c(2, 3), max), 1, w_stat)
  nW <- dim(W)[3]
  turn <- if (nW > 1) {
    dW <- abs(W[, , -1, drop = FALSE] - W[, , -nW, drop = FALSE])
    apply(apply(dW, c(2, 3), sum), 1, w_stat)
  } else {
    rep(NA_real_, length(res$strat))
  }
  est_names <- colnames(res$sumz)
  flips <- setNames(rep(NA_real_, length(res$strat)), res$strat)
  flips[paste0("TP_", est_names)] <- colSums(res$sumz[mask, , drop = FALSE] < 0)
  data.frame(
    AnnExRet = ann_ret, AnnVol = ann_vol, Sharpe = sharpe, CE = ce,
    GrossLev = gross, MaxWt = maxw, Turnover = turn,
    FlipWin = flips[res$strat], row.names = res$strat
  )
}
res_metrics <- oos_metrics(excess_returns, mean_estimators, cov_fun,
  training_weeks, testing_weeks, step_weeks,
  omega = omega
)
tab_full <- summarise_oos(res_metrics, masks$Full, omega = omega)
row_order <- c("EWP", paste0("MV_", est_order), paste0("TP_", est_order))
tab_full  <- tab_full[row_order, ]

# ========================================================================== #
# Debugging / tracability check: 1'Sigma^{-1}mu ~ 0
# > If the sum of unnormalised weights approaches zero the tangency portfolio
# > weights will explode to infinity (^ massive leverage).
# > This gets the worst-case window for the DMSh estimator to explore later
k_bad <- which.min(abs(res_metrics$sumz[, "DMSh"]))
sumz_bad <- res_metrics$sumz[k_bad, "DMSh"]
lev_bad <- sum(abs(res_metrics$W[, "TP_DMSh", k_bad]))
date_bad <- sr_df$TestStart[k_bad]
min_sumz <- apply(abs(res_metrics$sumz), 2, min)

# ========================================================================== #
# GICS (Global Industry Classification Standard) sector composition of
# sampled companies
id_col <- "DA441_column_name"
sector_col <- "GICS Sector Name"

sel_gics <- gics[gics[[id_col]] %in% selected, ]
sector_tab <- as.data.frame(table(sel_gics[[sector_col]]))
nrow(sel_gics) # check: should be 100 (Y)

colnames(sector_tab) <- c("Sector", "Count")
sector_tab <- sector_tab[order(-sector_tab$Count), ]

# BE note: quick calc to get number and % of GICS companies used in part A
# writeup
get_sec <- function(s) {
  v <- sector_tab$Count[sector_tab$Sector == s]
  if (length(v)) v else 0L
}
n_fin <- get_sec("Financials")
n_it <- get_sec("Information Technology")
n_comms <- get_sec("Communication Services")
pct_it_comms <- round(100 * (n_it + n_comms) / sum(sector_tab$Count))
```

```{r a-inout-compute, include=FALSE}
# In-sample vs out-of-sample tangency Sharpe, per window per estimator
in_out_sample <- function(X, dates, mean_estimators, cov_fun,
                          training_weeks, testing_weeks, step_weeks) {
  n_windows <- floor((nrow(X) - training_weeks - testing_weeks) / step_weeks) + 1
  rows <- list()
  i <- 1
  for (k in seq_len(n_windows)) {
    tr_start <- 1 + (k - 1) * step_weeks
    tr_end <- tr_start + training_weeks - 1
    te_start <- tr_end + 1
    te_end <- te_start + testing_weeks - 1
    train_X <- X[tr_start:tr_end, , drop = FALSE]
    test_X <- X[te_start:te_end, , drop = FALSE]
    Sigma_inv <- solve(cov_fun(train_X))
    for (nm in names(mean_estimators)) {
      z <- as.vector(Sigma_inv %*% mean_estimators[[nm]](train_X))
      w <- z / sum(z)
      rows[[i]] <- data.frame(
        window = k, Date = dates[te_start], estimator = nm,
        denom = sum(z),
        IS = ann_sharpe(train_X %*% w),
        OOS = ann_sharpe(test_X %*% w)
      )
      i <- i + 1
    }
  }
  out <- do.call(rbind, rows)
  out$flip <- out$denom < 0
  out
}
diag_df <- in_out_sample(
  excess_returns, dates, mean_estimators, cov_fun,
  training_weeks, testing_weeks, step_weeks
)
diag_df$estimator <- factor(diag_df$estimator, levels = est_order)
inout_summary <- diag_df |>
  dplyr::group_by(estimator) |>
  dplyr::summarise(
    `Windows flipped (%)` = mean(flip) * 100,
    `OOS Sharpe (flipped)` = mean(OOS[flip], na.rm = TRUE),
    `OOS Sharpe (not flipped)` = mean(OOS[!flip], na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::arrange(match(estimator, est_order))
```

---

## Data preprocessing {#a1}

The risk-free rate was first converted from a percentage to a decimal (i) before
weekly excess returns were computed (ii) using Eq. \@ref(eq:excess-returns).
\begin{equation}
r^{\text{excess}}_{i,t} = r_{i,t} - r_{f,t}
(\#eq:excess-returns)
\end{equation}
The risk-free rate was aligned to the return dates, and the excess returns were
calculated using the `sweep()` function to subtract the risk-free vector across
the rows. The resulting excess return matrix covers `r n_weeks` weekly
observations for `r n_assets` companies, from `r as.character(min(dates))` to 
`r as.character(max(dates))`. In **R**, both of these steps were implemented
using:

```{r a-data-prep-show, ref.label="a-data-prep", echo=TRUE, eval=FALSE}
```

## Selection of estimation models {#a2}

Portfolios were constructed using all eight mean-vector estimators: the sample 
mean, the four distribution-free shrinkage estimators of @Asimit15062026 
(`St-MSh`, `D-MSh`, `O-LSh` and `T-LSh`), and the three established benchmarks 
(`Wang`, `BOP` and `Jorion`). Please see Appendix
\@ref(apx-method-mean-vec-estimators) for formal definitions of each estimator.

Every mean estimator is paired with the one-parameter linear shrinkage estimator of @Ledoit2004365, `cov1Para`, rather than
the sample covariance. `cov1Para` pulls the sample covariance $S$ toward a
diagonal matrix $\mu I$ whose common variance
$\mu = \tfrac{1}{p}\operatorname{tr}(S)$ is the average sample variance. 
The strength of the pull is set automatically, with no tuning parameter, to minimise the covariance's mean-squared error (Appendix \@ref(apx-method-mean-vec-estimators)).
The result is always positive-definite and well-conditioned. It is also
distribution-free, needing only i.i.d. observations with finite fourth moments,
the same assumptions that underlie the mean estimators of @Asimit15062026. The
whole pipeline therefore rests on one consistent set of assumptions.

The issue with the sample covariance is conditioning. Each
training window gives $n = `r training_weeks`$ weekly observations on
$p = `r n_assets`$ assets, a ratio of $n/p \approx `r fmt(np_ratio, 2)`$.
Because $n > p$ the sample covariance can be inverted, but it is
ill-conditioned. Its smallest eigenvalues are biased toward zero, so inverting
the covariance amplifies the directions estimated least reliably
[@Ledoit2004365], producing extreme, unstable weights whose errors accumulate
out of sample.

Both the problem and the fix are visible directly in our data. Averaged across
the `r n_windows` windows (described in more detail in \@ref(a3)), the condition
number of the sample covariance is
`r fmt(cond_mean_sample, 0)` compared to `r fmt(cond_mean_shrink, 0)` under
`cov1Para` - a `r round(cond_ratio_mean)`-fold improvement (medians
`r fmt(cond_med_sample, 0)` vs `r fmt(cond_med_shrink, 0)`). 
With the mean estimators and portfolio rule held fixed, the better conditioning 
feeds straight into implementability. For the `O-LSh` tangency portfolio, 
average gross leverage falls from 
`r fmt(cond_tab["OLSh", "GrossLev_sample"], 1)` to
`r fmt(cond_tab["OLSh", "GrossLev_cov1Para"], 1)` and monthly turnover from
`r fmt(cond_tab["OLSh", "Turnover_sample"], 1)` to
`r fmt(cond_tab["OLSh", "Turnover_cov1Para"], 1)`, with comparable reductions
for every estimator (Appendix \@ref(apx-cond)). 

## Portfolio construction and rolling-window setup {#a3}

For each window and each mean estimator, two portfolios are built: a
mean-variance portfolio with risk aversion $\gamma = 1$, giving the unnormalised
weights $w_{MV}$, and a tangency portfolio, the same vector rescaled to sum to
one, $w_{TP}$,
$$
w_{MV} = \Sigma^{-1}\hat\mu \, ,
\qquad
w_{TP} = \frac{\Sigma^{-1}\hat\mu}{\mathbf 1^\top\Sigma^{-1}\hat\mu}
$$
The equal-weight portfolio (EWP) is included as a
benchmark. In our code, $\gamma$ is denoted `omega`. Throughout, $\Sigma$ is the
`cov1Para` estimate selected in \@ref(a2), and $\hat\mu$ is estimated on the
training window only, so no test-window information enters the weights.

The rolling structure uses a 156-week (3-year) training window, a 4-week
(1-month) testing window, and a 4-week step. Because the step equals the testing
length, successive test windows are contiguous and non-overlapping. Over the 
`r n_weeks`-week sample this gives 
$$
\lfloor(`r n_weeks` - 156 - 4)/4\rfloor + 1 = `r n_windows`
$$
out-of-sample windows, leaving the final 
`r n_weeks - 156 - testing_weeks * n_windows` weeks uncovered by any test window. The implementation of this rolling structure in **R** was:

```{=latex}
\newpage
```

```{r rolling-show, ref.label="a-rolling", echo=TRUE, eval=FALSE}
```


## Portfolio performance {#a4}

For every test window and every portfolio the annualised Sharpe ratio was
calculated by taking the mean of the four weekly out-of-sample excess returns,
dividing by their standard deviation and multiplying by $\sqrt{52}$. Each window
is assigned to a market environment by its test-window start date.
Of the `r n_windows` windows, `r n_fc` fall in the financial-crisis sub-period
(`r format(fc_lo, "%d/%m/%Y")`-`r format(fc_hi, "%d/%m/%Y")`) and `r n_pd` in
the pandemic sub-period 
(`r format(pd_lo, "%d/%m/%Y")`-`r format(pd_hi, "%d/%m/%Y")`).
These boundary dates are structural breakpoints identified by the Bai-Perron
test on the S&P500 return series and the sub-periods correspond to statistically
identified market regimes rather than arbitrary calendar choices. 

Table \@ref(tab:a-headline-sharpe) shows the mean and standard deviation of the
per-window Sharpe ratios over the full period and the two sub-periods. Each
per-window Sharpe ratio is estimated from only four weekly returns, so it is
inherently noisy. Cross-window standard deviations range from
approximately `r fmt(min(summary_full$SD), 1)` to 
`r fmt(max(summary_full$SD), 1)` over the full period, in 
every case several times the magnitude of the corresponding mean
(approximately `r fmt(min(summary_full$Mean), 2)` to 
`r fmt(max(summary_full$Mean), 2)`). This dispersion is a property of the
per-window definition. A companion summary for a Sharpe ratio computed on the
stitched out-of-sample series is described in Appendix \@ref(apx-scale).

Note that the Sample and `St-MSh` rows in Table \@ref(tab:a-headline-sharpe) are
identical in every period. This is a structural property of the Sharpe ratio. `St-MSh` scales the sample mean by a positive constant
that the Sharpe ratio cannot detect. This is explored further in Appendix
\@ref(apx-stmsh-vs-sample) (the same identity is also present in Table 3 of
@Asimit15062026).

```{r a-headline-sharpe}
hl <- combined_summary
hl$Strategy <- sub("^(MV|TP)_", "", hl$Strategy)
n_est <- length(est_order)

kable(hl,
  digits = 4, booktabs = TRUE, row.names = FALSE,
  caption = "Mean and standard deviation of the per-window annualised Sharpe ratios. The best mean and standard deviation in each environment are highlighted. Because the Sharpe ratio is invariant to positive rescaling of the weights, St-MSh reproduces the Sample rows exactly, and the mean-variance and tangency portfolios coincide except in net-short windows where the sign flips.",
  linesep = "",
  col.names = c("Strategy", "Mean", "SD", "Mean", "SD", "Mean", "SD")
) |>
  kable_styling(
    full_width = FALSE, position = "center", 
    latex_options = "hold_position"
  ) |>
  add_header_above(c(" " = 1, "Full period" = 2, "Financial crisis" = 2, "Pandemic" = 2)) |>
  pack_rows(index = c(
    "Benchmark" = 1,
    "Mean-variance portfolio" = n_est,
    "Tangency portfolio" = n_est
  )) |>
  # adds latex spacing (just for pdf)
  row_spec(c(5, 13), extra_latex_after = "\\addlinespace") |>
  column_spec(2,
    bold = combined_summary$Mean_full == max(combined_summary$Mean_full, na.rm = TRUE),
    color = ifelse(combined_summary$Mean_full == max(combined_summary$Mean_full, na.rm = TRUE), "#1b9e77", "black")
  ) |>
  column_spec(4,
    bold = combined_summary$Mean_fc == max(combined_summary$Mean_fc, na.rm = TRUE),
    color = ifelse(combined_summary$Mean_fc == max(combined_summary$Mean_fc, na.rm = TRUE), "#1b9e77", "black")
  ) |>
  column_spec(6,
    bold = combined_summary$Mean_pd == max(combined_summary$Mean_pd, na.rm = TRUE),
    color = ifelse(combined_summary$Mean_pd == max(combined_summary$Mean_pd, na.rm = TRUE), "#1b9e77", "black")
  ) |>
  column_spec(3,
    bold = combined_summary$SD_full == min(combined_summary$SD_full, na.rm = TRUE),
    color = ifelse(combined_summary$SD_full == min(combined_summary$SD_full, na.rm = TRUE), "#1b9e77", "black")
  ) |>
  column_spec(5,
    bold = combined_summary$SD_fc == min(combined_summary$SD_fc, na.rm = TRUE),
    color = ifelse(combined_summary$SD_fc == min(combined_summary$SD_fc, na.rm = TRUE), "#1b9e77", "black")
  ) |>
  column_spec(7,
    bold = combined_summary$SD_pd == min(combined_summary$SD_pd, na.rm = TRUE),
    color = ifelse(combined_summary$SD_pd == min(combined_summary$SD_pd, na.rm = TRUE), "#1b9e77", "black")
  )
```


## Interpretation of results and final report {#a5}

<!-- BE note: was quite brutal here to get this down -->
<!-- - Aimed for ~1/2 to 1 page max -->

Optimised portfolios are only as reliable as the expected returns they are built
on, and those estimates carry substantial estimation error. This report tests
whether shrinkage methods designed to temper that noise deliver better
risk-adjusted returns than the classical sample mean for an investor rebalancing
monthly over two decades [@Asimit15062026]. Using weekly excess returns for
`r no_companies` S&P 500 companies 
(`r format(min(dates), "%B %Y")`-`r format(max(dates), "%B %Y")`), 
eight ways of estimating expected returns were compared, each paired with the
`cov1Para` covariance estimator [@Ledoit2004365] and benchmarked against an
equal-weight (1/N) portfolio. Every method was evaluated with a rolling window
that mirrors live practice, fiting on the previous three years, testing on the
next unseen month, then advancing, giving `r n_windows` out-of-sample tests with
no look-ahead bias. Performance was summarised by the annualised Sharpe ratio
over the full period, the 2008 financial crisis, and the COVID-19 pandemic
(Table \@ref(tab:a-headline-sharpe)).

The mean-variance and tangency portfolios hold the same positions at different
scales (tangency rescales the weights to sum to one) so within any window
their Sharpe ratios are identical bar a possible sign flip (the per-window
magnitudes agree to within `r formatC(mvtp_maxdiff, format = "e", digits = 1)`).
The two coincide in every window where the portfolio is not net-short, which
covers almost the entire pandemic. Over other periods the sign flip lifts the
tangency full-period mean above mean-variance for the stronger estimators
(`TP_OLSh`
`r fmt(summary_full["TP_OLSh", "Mean"], 4)` versus `MV_OLSh`
`r fmt(summary_full["MV_OLSh", "Mean"], 4)`). The economically meaningful
difference is cost, not Sharpe ratio. Because tangency weights sum to one,
leverage and turnover are far lower (for `O-LSh`, average leverage falls from
`r fmt(tab_full["MV_OLSh", "GrossLev"], 1)` to
`r fmt(tab_full["TP_OLSh", "GrossLev"], 1)`), making tangency preferable in
practice. The exception is when the normaliser nears zero and weights explode
(`r format(date_bad, "%d %B %Y")` drives `TP_DMSh` leverage to roughly
`r fmt(lev_bad, 0)`; Appendix \@ref(apx-pathology)).

Across estimators, the standout result is how hard the naïve benchmark is to
beat. The equal-weight portfolio posts the highest full-period Sharpe
(`r fmt(summary_full["EWP", "Mean"], 4)`) and lowest full-period dispersion,
overtaken only once, by `TP_OLSh` in the financial crisis
(`r fmt(summary_fc["TP_OLSh", "Mean"], 4)` versus
`r fmt(summary_fc["EWP", "Mean"], 4)`), consistent with the 1/N result of
@DeMiguel2009. Among optimised methods only mean shrinkage keeps up. Under the tangency construction, `O-LSh` and `D-MSh` roughly double the sample mean's full-period Sharpe ratio 
(`r fmt(summary_full["TP_OLSh", "Mean"], 4)` and
`r fmt(summary_full["TP_DMSh", "Mean"], 4)` versus 
`r fmt(summary_full["TP_Sample", "Mean"], 4)`), with `D-MSh` the most consistent across regimes and the best optimised pandemic performer 
(`r fmt(summary_pd["TP_DMSh", "Mean"], 4)`). 
The raw sample mean turns negative in the financial crisis 
(`r fmt(summary_fc["MV_Sample", "Mean"], 4)`) through error-maximisation,
`St-MSh` reproduces the sample rows exactly (Appendix
\@ref(apx-stmsh-vs-sample)), `Jorion` is too volatile to trust
(SD `r fmt(summary_full["MV_Jorion", "SD"], 2)`, rising to
`r fmt(summary_pd["TP_Jorion", "SD"], 2)` in the pandemic), and `BOP` is the
weakest overall, with the most net-short flips (`r tab_full["TP_BOP",
"FlipWin"]` windows). The uniformly weak pandemic figures are partly due to our
group's sector draw (Figure \@ref(fig:gics-fig); full company list in Appendix
\@ref(apx-stocks)): Financials is the largest single exposure (`r n_fin` of
100), and while Information Technology sits mid-pack (`r n_it`), the tech-led
sectors that drove the post-2020 rally are thin as a bloc with IT and
Communication Services (`r n_comms`) together only accounting for 
`r pct_it_comms`% of the universe. No estimator can manufacture exposure that 
isn't there and an equal-weighted portfolio sidesteps this problem by never guessing. 
For our 100 stocks, `O-LSh` and `D-MSh` are worth using over the raw sample
mean, but once leverage and turnover are counted, a plain equal-weighted
portfolio remains a strong, cheap benchmark the optimisers struggle to beat out
of sample (Appendix \@ref(apx-scale)).


```{r gics-table, eval=FALSE}
# BE note: not loaded as opted for the bar chart below instead
# Kept for posterity
kable(sector_tab,
  row.names = FALSE, booktabs = TRUE,
  caption = "GICS sector composition of our 100 sampled companies (Group 12)."
) |>
  kable_styling(full_width = FALSE, position = "center")
```

```{r gics-fig, fig.cap="GICS sector composition of the 100 sampled companies (Group 12), shown as a horizontal bar chart with sectors ordered by company count. Financials is the largest exposure, while Communication Services and the wider tech-led bloc are comparatively underweight.", fig.width=5.5, fig.height=2.8}
ggplot(sector_tab, aes(x = Count, y = reorder(Sector, Count))) +
  geom_col(fill = "#0072B2") +
  labs(x = "Companies", y = NULL) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, by = 5)) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )
```

```{=latex}
\newpage
```

# Coursework 2 {-}

# Shrinkage Generalised Linear Models for Insurance Claim Severity Estimation

```{=latex}
% Custom shortened header to avoid overlap
\sectionmark{Shrinkage GLMs for Insurance Claim Severity Estimation}
```

---


<!-- # Conclusions

TBI -->

<!-- ```{=html} -->
<!-- # References -->
<!-- <div id="refs"></div> -->
<!-- ``` -->

<!-- ```{=latex} -->
<!-- \bibliography{references.bib} -->
<!-- ``` -->

```{=latex}
\clearpage
\newpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
```

# (APPENDIX) Appendices {-}

```{=latex}
\titleformat{\section}[block]{\LARGE\bfseries}{\thesection}{1em}{}
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}
\setcounter{section}{0}
```

# Coursework 1: Supplemental methodology {#apd-supplementary-methods}

## Mean-vector estimators {#apx-method-mean-vec-estimators}

Every estimator starts from the sample mean vector, $\bar X$. The shrinkage
estimators adjust $\bar X$ using the unbiased sample covariance of the returns,
$S$, and the grand mean $\bar X_g = \tfrac{1}{p}\mathbf 1^\top \bar X$, the
average of $\bar X$ across the $p$ assets. Each estimator maps the $n \times p$
matrix of training returns to its own length-$p$ estimate, and the estimators
are compared by mean squared error (MSE).

```{asis, echo=knitr::is_html_output()}
| Estimator                                                   | Shrinks toward                | How it shrinks                                             | Loss      | Source          |
| ----------------------------------------------------------- | ----------------------------- | ---------------------------------------------------------- | --------- | --------------- |
| Sample mean                                                 | no shrinkage                  | Unbiased. High variance when $p \approx n$                 | -         | -               |
| Stein Multiplicative Shrinkage `St-MSh`                     | zero                          | One scalar $\hat c \in [0,1)$ applied to every asset       | MSE       | @Asimit15062026 |
| Diagonal Multiplicative Shrinkage `D-MSh`                   | zero                          | Asset-specific scalars $\hat c_k$ (suits differing scales) | MSE       | @Asimit15062026 |
| One-Parameter Linear Shrinkage `O-LSh`                      | grand mean                    | One intensity $\hat\delta$                                 | MSE       | @Asimit15062026 |
| Two-Parameter Linear Shrinkage `T-LSh`                      | grand mean, rescaled          | Intensity $\hat\delta$ plus target scaling $\hat\xi$       | MSE       | @Asimit15062026 |
| The shrinkage estimator in @Wang2014 `Wang`                 | grand mean                    | One intensity, estimated by U-statistics                   | MSE       | @Wang2014       |
| The shrinkage estimator in @Bodnar2019 `BOP`                | precision-weighted grand mean | One intensity, built from $S^{+}$                          | invariant | @Bodnar2019     |
| The Bayes-Stein shrinkage estimator in @Jorion1986 `Jorion` | precision-weighted grand mean | Bayes-Stein intensity                                      | invariant | @Jorion1986     |
```

```{r apx-mean-estimator-table, eval=knitr::is_latex_output()}
est_tab <- data.frame(
  Estimator = c(
    "Sample mean",
    "Stein Multiplicative Shrinkage (\\greentt{St-MSh})",
    "Diagonal Multiplicative Shrinkage (\\greentt{D-MSh})",
    "One-Parameter Linear Shrinkage (\\greentt{O-LSh})",
    "Two-Parameter Linear Shrinkage (\\greentt{T-LSh})",
    "Shrinkage estimator of Wang et al. (\\greentt{Wang})",
    "Shrinkage estimator of Bodnar et al. (\\greentt{BOP})",
    "Bayes--Stein estimator (\\greentt{Jorion})"
  ),
  Target = c(
    "no shrinkage",
    "zero",
    "zero",
    "grand mean",
    "grand mean, rescaled",
    "grand mean",
    "precision-weighted grand mean",
    "precision-weighted grand mean"
  ),
  How = c(
    "Unbiased; high variance when $p \\approx n$",
    "One scalar $\\hat c \\in [0,1)$ applied to every asset",
    "Asset-specific scalars $\\hat c_k$ (suits differing scales)",
    "One intensity $\\hat\\delta$",
    "Intensity $\\hat\\delta$ plus target scaling $\\hat\\xi$",
    "One intensity, estimated by U-statistics",
    "One intensity, built from $S^{+}$",
    "Bayes--Stein intensity"
  ),
  Loss = c(
    "---", "MSE", "MSE", "MSE", "MSE", "MSE",
    "invariant", "invariant"
  ),
  # manually coded
  Source = c( 
    "---",
    "\\citet{Asimit15062026}", "\\citet{Asimit15062026}",
    "\\citet{Asimit15062026}", "\\citet{Asimit15062026}",
    "\\citet{Wang2014}", "\\citet{Bodnar2019}", "\\citet{Jorion1986}"
  )
)
kable(est_tab,
  booktabs = TRUE, escape = FALSE, row.names = FALSE,
  caption = "Mean-vector estimators: shrinkage target, mechanism, and the loss each is designed to minimise.",
  col.names = c("Estimator", "Shrinks toward", "How it shrinks", "Loss", "Source"),
  linesep = ""
) |>
  kable_styling(
    full_width = FALSE, position = "center",
    font_size = 9, latex_options = "hold_position"
  ) |>
  column_spec(1, width = "13em") |>
  column_spec(2, width = "8em") |>
  column_spec(3, width = "13em")
```


The four estimators of @Asimit15062026 take the following explicit forms:

$$
\hat\mu^{\text{St-MSh}} = \hat c\,\bar X,\qquad
\hat c = \frac{\|\bar X\|_p^2}{\|\bar X\|_p^2 + \tfrac{1}{np}\operatorname{tr}(S)},
\qquad \|\bar X\|_p^2 = \tfrac{1}{p}\bar X^\top \bar X,
$$
$$
\hat\mu^{\text{D-MSh}}_k = \hat c_k\,\bar X_k,\qquad
\hat c_k = \frac{\bar X_k^2}{\bar X_k^2 + \tfrac{1}{n}S_{kk}},
$$
$$
\hat\mu^{\text{O-LSh}} = (1-\hat\delta)\bar X + \hat\delta\,\bar X_g\mathbf 1,
\qquad
\hat\mu^{\text{T-LSh}} = (1-\hat\delta)\bar X + \hat\delta\hat\xi\,\bar X_g\mathbf 1.
$$

The amount of shrinkage in each estimator is set by a shrinkage intensity, 
$\hat c$ or $\hat c_k$. This scalar multiplies the sample mean in `St-MSh` and
`D-MSh`, or the weight $\hat\delta$ that mixes the sample mean with the target
in `O-LSh` and `T-LSh`. All four intensities follow the same signal-to-noise
trade-off. They shrink harder when the sample mean is noisier and less when the
signal is stronger. `T-LSh` extends `O-LSh` by also scaling the target, and its
oracle MSE (the MSE it would attain if the true mean and covariance were known)
is strictly lower than that of `O-LSh`.

The other three estimators, `Wang`, `BOP` and `Jorion`, are established
benchmarks. `Wang` shares the target of `O-LSh`, the grand mean, but estimates
its intensity with U-statistics rather than the sample plug-in estimates used
above [@Wang2014]. `BOP` and `Jorion` instead shrink toward a precision-weighted
grand mean, a minimum-variance combination formed with $S^{+}$ for `BOP` and
$\tilde S^{-1}$ for `Jorion` [@Bodnar2019; @Jorion1986]. Both optimise the
invariant loss rather than the MSE, weighting estimation errors by the precision
matrix, the inverse covariance.

## Why `St-MSh` reproduces the sample-mean Sharpe ratios exactly {#apx-stmsh-vs-sample}

The identical Sample and `St-MSh` rows in Table \@ref(tab:a-headline-sharpe) are
not a coincidence, nor a bug. The same identity appears in Table 3 of
@Asimit15062026 where the `St-MSh` and Sample rows are identical in both the
mean-variance and tangency panels (e.g. 0.274/1.352 in Panel A and 0.385/1.325
in Panel B), despite their different universe (441 stocks), frequency (daily)
and window scheme.

This occurs because `St-MSh` multiplies the sample mean by a single scalar, 
$\hat\mu^{\text{St-MSh}} = \hat c\,\bar X$, and this scalar is strictly
positive, since the numerator of $\hat c$ is a squared norm and the denominator
adds a sum of sample variances. Because the portfolio map is linear in
$\hat\mu$, the scalar passes straight through to the weights,
$$
w_{MV}^{\text{St-MSh}}
  = \Sigma^{-1}(\hat c\,\bar X)
  = \hat c\, w_{MV}^{\text{Sample}},
$$
and a portfolio's weekly returns scale with its weights,
$r_t(\hat c\,w) = \hat c\, r_t(w)$. The Sharpe ratio is a ratio of a mean to a
standard deviation, so any positive constant cancels:
$$
\mathrm{SR}(\hat c\,w)
  = \frac{\hat c\,\bar r}{\hat c\, s_r}\sqrt{52}
  = \mathrm{SR}(w).
$$
The two strategies therefore have the same Sharpe ratio in every window, hence
the same mean and the same standard deviation across windows. 
<!-- Note to Vali: this is verified numerically: `r stm_check` -->
For the tangency portfolio the cancellation happens one step earlier, in the
weights themselves,
$$
w_{TP}^{\text{St-MSh}}
  = \frac{\hat c\,\Sigma^{-1}\bar X}{\hat c\,\mathbf 1^\top\Sigma^{-1}\bar X}
  = w_{TP}^{\text{Sample}},
$$
As a result `TP_StMSh` and `TP_Sample` are the same portfolio with identical leverage, turnover and net-short windows, as seen in Tables \@ref(tab:cond-table) and
\@ref(tab:scale-metrics). Since $\hat c > 0$, `St-MSh` also flips sign in exactly the same windows as the sample mean.

An interesting observation is that because the scalar $\hat c$ is
re-estimated each window the stitched `MV_StMSh` series is not a global
rescaling of `MV_Sample` as each month is scaled by a different constant. The
per-window Sharpe statistics are identical, but the stitched Sharpe in Table
\@ref(tab:scale-metrics) differs slightly, and every scale-sensitive metric
(leverage, volatility, certainty-equivalent) shrinks by roughly $\hat c$ for the
mean-variance portfolio. The other estimators avoid this equivalence
because they change the direction. `D-MSh` rescales each asset differently, and
the linear-shrinkage estimators mix $\bar X$ with a nonzero target. `St-MSh`'s
shrinkage is visible in the scale-sensitive mean-variance metrics, but it is
invisible to any scale-invariant evaluation of a linear portfolio rule.

## Covariance conditioning: `cov1Para` vs sample covariance {#apx-cond}

With the mean estimators and portfolio construction held fixed, replacing
`cov1Para` by the sample covariance sharply worsens conditioning and
implementability. Averaged across windows, the condition number of the
covariance estimate is
`r fmt(cond_mean_shrink, 1)` under `cov1Para` against
`r fmt(cond_mean_sample, 1)` under the sample covariance, with medians of
`r fmt(cond_med_shrink, 1)` and `r fmt(cond_med_sample, 1)`. Table
\@ref(tab:cond-table) shows, for each mean estimator, the average gross
leverage and turnover of the tangency portfolio under each covariance estimator,
along with its Sharpe ratio on the stitched out-of-sample series.

```{r cond-table}
kbl_cond <- cond_tab[, c(
  "GrossLev_cov1Para", "Turnover_cov1Para", "OOS_Sharpe_cov1Para",
  "GrossLev_sample", "Turnover_sample", "OOS_Sharpe_sample"
)]
kable(kbl_cond,
  digits = 2, booktabs = TRUE,
  caption = "Tangency-portfolio implementability by covariance estimator. Gross leverage and turnover are averaged across windows, and Sharpe is computed on the stitched out-of-sample series.",
  col.names = c("Gross lev.", "Turnover", "Sharpe", "Gross lev.", "Turnover", "Sharpe")
) |>
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") |>
  add_header_above(c(" " = 1, "cov1Para" = 3, "Sample covariance" = 3))
```

```{=latex}
\newpage
```

## Scale-sensitive out-of-sample metrics {#apx-scale}

<!-- The annualised Sharpe ratio is invariant to positive rescaling of the weight -->
<!-- vector. It cannot distinguish -->
<!-- portfolios that are far apart in leverage, trading, and exposure, differences -->
<!-- that determine whether a strategy can actually be run. The metrics in -->
<!-- Table \@ref(tab:scale-metrics) are chosen because they respond to scale and they are computed on the stitched out-of-sample series, -->
<!-- with the test windows concatenated in time order. This is why the Sharpe column -->
<!-- differs from the per-window average in Table \@ref(tab:a-headline-sharpe). CE is -->
<!-- the realised certainty-equivalent return, -->
<!-- $\text{annual return} - \tfrac{\omega}{2}\,\text{annual variance}$, where -->
<!-- $\omega = 1$ is the evaluation risk aversion, distinct from the $\gamma$ that -->
<!-- scales the portfolio weights. Leverage and max weight are averaged across -->
<!-- windows, turnover is per monthly rebalance, and net-short counts the windows with -->
<!-- $c = \mathbf 1^\top\Sigma^{-1}\hat\mu < 0$. -->



The annualised Sharpe ratio is invariant to positive rescaling of the weight
vector, so it cannot see differences that matter for implementation. The metrics
in Table \@ref(tab:scale-metrics) are not scale-invariant: they are computed on
the stitched out-of-sample series (test windows concatenated in time order), so
the Sharpe column differs from the per-window average in
Table \@ref(tab:a-headline-sharpe). CE is the realised certainty-equivalent return,
$\text{annual return} - \tfrac{\omega}{2}\,\text{annual variance}$ with
$\omega = 1$; leverage and max weight are averaged across windows; turnover is per
monthly rebalance; net-short counts windows with
$c = \mathbf 1^\top\Sigma^{-1}\hat\mu < 0$.

Note: Turnover compares target weights at consecutive rebalances and does not account for the drift of holdings by intervening returns. It therefore approximates true trading volume, and because the same simplification applies to every strategy, cross-strategy comparisons are unaffected.

```{r scale-metrics}
# options(knitr.kable.NA = "---")
# best_sr <- tab_full$Sharpe == max(tab_full$Sharpe, na.rm = TRUE)
# best_ce <- tab_full$CE == max(tab_full$CE, na.rm = TRUE)
# kable(tab_full,
#   digits = c(2, 2, 2, 1, 1, 2, 1, 0),
#   format.args = list(big.mark = ","), booktabs = TRUE,
#   caption = "Scale-sensitive out-of-sample metrics, full period (annualised weekly excess returns, omega = 1). Sharpe is on the stitched OOS series; CE is the realised certainty-equivalent; leverage and max weight are window averages; turnover is per monthly rebalance; net-short counts windows with negative sum(z).",
#   col.names = c(
#     "Ann. ex. ret.", "Ann. vol.", "Sharpe", "CE",
#     "Gross lev.", "Max wt", "Turnover", "Net-short"
#   ),
#   linesep = c("\\addlinespace", rep(c("", "\\addlinespace"), length(mean_estimators)))
# ) |>
#   kable_styling(
#     full_width = FALSE, position = "center", font_size = 9,
#     latex_options = c("scale_down", "hold_position")
#   ) |>
#   add_header_above(c(" " = 1, "Performance" = 4, "Implementability" = 4)) |>
#   column_spec(3, bold = best_sr, color = ifelse(best_sr, "#1b9e77", "black")) |>
#   column_spec(4, bold = best_ce, color = ifelse(best_ce, "#1b9e77", "black"))
options(knitr.kable.NA = "---")

tf <- data.frame(
  Strategy = sub("^(MV|TP)_", "", rownames(tab_full)),
  tab_full, row.names = NULL, check.names = FALSE
)

best_sr <- tf$Sharpe == max(tf$Sharpe, na.rm = TRUE)
best_ce <- tf$CE     == max(tf$CE,     na.rm = TRUE)

kable(tf,
  digits = c(0, 2, 2, 2, 1, 1, 2, 1, 0),   # leading 0 for Strategy
  format.args = list(big.mark = ","), booktabs = TRUE, row.names = FALSE,
  caption = "Scale-sensitive out-of-sample metrics, full period (annualised weekly excess returns, omega = 1). Sharpe is on the stitched OOS series; CE is the realised certainty-equivalent; leverage and max weight are window averages; turnover is per monthly rebalance; net-short counts windows with negative sum(z).",
  col.names = c(
    "Strategy", "Ann. ex. ret.", "Ann. vol.", "Sharpe", "CE",
    "Gross lev.", "Max wt", "Turnover", "Net-short"
  ),
  linesep = ""
) |>
  kable_styling(
    full_width = FALSE, position = "center", font_size = 9,
    latex_options = c("scale_down", "hold_position")
  ) |>
  add_header_above(c(" " = 1, "Performance" = 4, "Implementability" = 4)) |>
  pack_rows(index = c(
    "Benchmark" = 1,
    "Mean-variance portfolio" = length(est_order),
    "Tangency portfolio" = length(est_order)
  )) |>
  column_spec(4, bold = best_sr, color = ifelse(best_sr, "#1b9e77", "black")) |>
  column_spec(5, bold = best_ce, color = ifelse(best_ce, "#1b9e77", "black"))
```

## In-sample versus out-of-sample Sharpe {#apx-inout}

The tangency weights are $\Sigma^{-1}\mu / (\mathbf 1^\top\Sigma^{-1}\mu)$; when
the denominator $c = \mathbf 1^\top\Sigma^{-1}\mu$ is negative, the "maximiser" in
fact returns the minimum-Sharpe portfolio, so the in-sample Sharpe goes large and
negative. Counting those sign-flip windows is a useful diagnostic - they cluster
in high-volatility regimes, exactly where the equal-weight portfolio wins because
it never optimises. Table \@ref(tab:inout-table) reports, per estimator, the share
of windows that flip and the average out-of-sample Sharpe in flipped versus
non-flipped windows; Figure \@ref(fig:inout-fig) plots in- against out-of-sample
Sharpe.

As elsewhere, the Sample and `St-MSh` rows coincide exactly: since
$\hat c > 0$, the sign of $c = \mathbf 1^\top\Sigma^{-1}\hat\mu$ is unchanged
by the shrinkage, so the two estimators flip in precisely the same windows
(Appendix \@ref(apx-stmsh-vs-sample)).

```{r inout-table}
kable(inout_summary,
  digits = 2, booktabs = TRUE, 
  caption = "Tangency sign-flip diagnostic by estimator: share of net-short windows and average out-of-sample Sharpe in flipped versus non-flipped windows.",
  col.names = c(
    "Estimator", "Windows flipped (\\%)",
    "OOS Sharpe (flipped)", "OOS Sharpe (not flipped)"
  ),
  escape = FALSE
) |>
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position")
```

```{r inout-fig, fig.cap="In-sample versus out-of-sample annualised Sharpe (tangency portfolios), by estimator. Net-short windows (sign flips) are highlighted.", fig.width=6.5, fig.height=4.5}
ggplot(diag_df, aes(IS, OOS, color = flip)) +
  geom_hline(yintercept = 0, linewidth = 0.3) +
  geom_vline(xintercept = 0, linewidth = 0.3) +
  geom_point(alpha = 0.5, size = 1) +
  facet_wrap(~estimator) +
  labs(x = "In-sample Sharpe", y = "Out-of-sample Sharpe", color = "Sign flip") +
  theme_minimal()
```

## The $\mathbf 1^\top\Sigma^{-1}\hat\mu \approx 0$ pathology {#apx-pathology}

Tangency gross leverage in window $k$ is $\sum_i |z_i| / |c_k|$ with
$c_k = \mathbf 1^\top\Sigma^{-1}\hat\mu$. When a window lands with $c_k$ close to
zero, the weights explode and every moment-based metric (return, volatility, CE,
leverage, turnover) is dominated by that one window - while the window's
per-window Sharpe is unchanged (it is scale-invariant), which is why the headline
table cannot see the problem. The closest approach to zero across windows is
smallest for D-MSh: at window `r k_bad` (`r format(date_bad, "%d %B %Y")`),
$c \approx `r formatC(sumz_bad, format = "f", digits = 4)`$ drives TP_DMSh gross
leverage to about `r fmt(lev_bad, 0)`. Table \@ref(tab:pathology-table) lists the
minimum $|c|$ per estimator.

```{r pathology-table}
patho <- data.frame(
  Estimator = est_order,
  `Min |c|` = round(as.numeric(min_sumz[est_order]), 4),
  check.names = FALSE
)
kable(patho,
  row.names = FALSE, booktabs = TRUE,
  caption = "Closest approach of $c = \\mathbf{1}^\\top \\Sigma^{-1}\\hat\\mu$ to zero, by estimator (smaller means a more unstable tangency window)."
) |>
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position")
```

## Sampled companies and sector exposure {#apx-stocks}

Figure \@ref(fig:gics-fig) (in \@ref(a5)) shows the sector composition. The full
list of the 100 sampled companies, grouped by GICS sector, is in Table
\@ref(tab:stocks-table).

```{r stocks-table}
stocks_tbl <- gics[
  gics$DA441_column_name %in% selected,
  c("company_ticker", "company_name", "GICS Sector Name")
]
stocks_tbl <- stocks_tbl[order(stocks_tbl[["GICS Sector Name"]], stocks_tbl$company_ticker), ]
colnames(stocks_tbl) <- c("Ticker", "Company", "GICS Sector")
idx <- rle(as.character(stocks_tbl$`GICS Sector`)) # sector run-lengths (table is sector-sorted)
kable(stocks_tbl[, c("Ticker", "Company")],
  booktabs = TRUE, longtable = TRUE,
  caption = "Our (Group 12's) 100 sampled companies, grouped by GICS sector."
) |>
  kable_styling(full_width = FALSE, font_size = 9, latex_options = c("repeat_header")) |>
  pack_rows(index = setNames(idx$lengths, idx$values))
```

## Our `r no_companies` stocks (Group `r group_num`) versus the full universe {#apx-universe}

Raw returns, equal-weighted, on a log axis so the 2008 and 2020 drawdowns stay
legible across 24 years. The full-441 equal-weight series is an equal-weight proxy
for the market, not the cap-weighted S&P 500. Figure \@ref(fig:universe-fig)
compares the Group 12 sample with the full universe; Figure \@ref(fig:sector-fig)
shows the cumulative path of each GICS sector, illustrating why the sample's light
Technology weighting muted its pandemic Sharpe ratios.

```{r apx-a-stock-plots, fig.cap="Rebased equal-weighted values for the SP 500 from x to y!. Top: Group 12 sample versus the full X*! universe. Bottom: Full XJ*! universe grouped by GICS sector.", fig.width=6.5, fig.height=7.5}
regimes <- data.frame(
  Regime = c("Financial crisis", "Pandemic"),
  xmin = c(fc_lo, pd_lo),
  xmax = c(fc_hi, pd_hi)
)

# BE note: don't like hard coded values here
cum_df <- data.frame(
  Date = dates,
  `Group 12 (100)` = cumprod(1 + rowMeans(returns100)),
  `All DA441 (441)` = cumprod(1 + rowMeans(returns_all)),
  check.names = FALSE
)

long_cum <- pivot_longer(
  cum_df, -Date,
  names_to = "Universe", values_to = "Growth"
)
ylims <- range(long_cum$Growth, na.rm = TRUE)

p1 <- ggplot(
  pivot_longer(cum_df, -Date, names_to = "Universe", values_to = "Growth"),
  aes(Date, Growth, color = Universe)
) +
  geom_rect(
    data = regimes,
    inherit.aes = FALSE,
    # manually tweaked
    aes(
      xmin = xmin, xmax = xmax, ymin = ylims[1], ymax = ylims[2], fill = Regime
    ),
    alpha = 0.08
  ) +
  scale_fill_manual(
    breaks = c("Financial crisis", "Pandemic"),
    values = c("Financial crisis" = "red", "Pandemic" = "blue")
  ) +
  geom_line() +
  scale_y_log10() +
  labs(y = "Rebased values (log $)", x = NULL) +
  theme_minimal() +
  theme(legend.position = "top", legend.title = element_blank()) +
  guides(color = guide_legend(order = 1), fill = guide_legend(order = 2))

sec_map <- setNames(gics[["GICS Sector Name"]], gics$DA441_column_name)
sectors <- sec_map[colnames(returns_all)]
stopifnot(!anyNA(sectors))

sect_cum <- sapply(sort(unique(sectors)), function(s) {
  cumprod(1 + rowMeans(returns_all[, sectors == s, drop = FALSE]))
})

sect_df <- data.frame(Date = dates, sect_cum, check.names = FALSE)

p2 <- ggplot(
  pivot_longer(sect_df, -Date, names_to = "Sector", values_to = "Growth"),
  aes(Date, Growth, color = Sector)
) +
  geom_line(linewidth = 0.5) +
  scale_y_log10() +
  labs(y = "Rebased values (log $)", x = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  guides(color = guide_legend(nrow = 3))

p1 / p2
```

```{=latex}
\clearpage
\newpage
```

# Coursework 2: supplementary analyses {#apd-b}



# Reproducibility, accessibility & AI declaration

<!-- BE note: taken out as we are submitting R file separately -->
<!-- Use: `knitr::purl("cs1-group07.rmd", documentation = 0)` to generate r file -->
<!-- Use: `tools::showNonASCIIfile("cs1-group07.rmd")` to debug -->
<!-- ## R Code Documentation -->
<!-- ```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}``` -->

## Reproducibility & accessibility

An accessible HTML version of this report is available via a public GitHub page:

- !Link TBI

This report was created in R Markdown. The source code is open source and is
available via:

- !Link TBI

Changes made to this document were tracked using Git and are also available via
the same repository (!link TBI). Please note that this commit history reflects
updates made to the project repository and is not a comprehensive timeline of
all contributions by group members.

## Generative AI Declaration

Generative AI (GenAI) tools were used throughout this project to assist with
project planning, methodological validation, and debugging. In practice, these
tools functioned as an additional technical team member, providing immediate
feedback, troubleshooting code, and challenging statistical assumptions.

### Tools Used

! To be updated... Google Gemini 3 Pro (Deep Research), ChatGPT 5.2

<!-- ### Prompts used -->

<!-- The GenAI tools were provided with the 'coursework_CM2_250120.pdf' file and -->
<!-- extracts from our `R` code. -->

<!-- A non-exhaustive list of prompts used is shown below. We opted to include at -->
<!-- least one example of each _type_ of prompt used, rather than a list of all -->
<!-- prompts. For example, if we entered the prompts 'How can we put two ggplots side -->
<!-- by side?' and 'How to remove grey background on a ggplot', we would include only -->
<!-- the first of these in the list below, as both relate to plotting issues (& are -->
<!-- similar to 'querying' with a traditional search engine). -->

<!-- - "...the top of both graphs in r are slightly different. (Pasting R code.) Can -->
<!--   this be fixed for the plot area?" -->
<!-- - "Help to identify the key points here" -->
<!-- - "How can this error be tracked down in an rmd file? ! Text line contains an -->
<!--   invalid character. l.1 ^^@^^@..." -->

<!-- ### Use of GenAI outputs & subsequent revisions -->

<!-- The GenAI tools served as a "consultant" member of the group. Notably, the AI -->
<!-- output: -->

<!-- - Helped debug plotting issues. -->
<!-- - Commented on the structure of R code snippets and the report structure. -->

<!-- ### Changes made: -->

<!-- All AI suggestions were critically reviewed and debated by the group before -->
<!-- implementation. For example: -->

<!-- - While the AI suggested code for complex layouts, the specific aesthetic -->
<!--   choices and annotations were manually adjusted to ensure clarity. -->
<!-- - Where the AI proposed specific interpretations, these were cross-referenced -->
<!--   with academic literature and statistics textbooks. -->
<!-- - The final R code for implementation was written and debugged by the group with -->
<!--   AI snippets being treated as drafts rather than final solutions. -->

<!-- While GenAI was treated as a collaborator, the final analysis, statistical -->
<!-- interpretation, and the written content of this report were generated and -->
<!-- verified by the (human) group members. -->

<!-- [^utility-theory]: -->
<!--     Described in more detail in the appendix (\S\@ref(apd-utility-theory)). -->

<!-- [^utility-skew]: -->
<!--     This is demonstrated in the appendix (\S\@ref(apd-utility-theory) and Table -->
<!--     \@ref(tab:exact-utility-table)) where we show that a constant relative risk -->
<!--     aversion (CRRA) utility function with $\gamma > 0$ mathematically confirms -->
<!--     that for risk-averse investors asset 1 results in a lower expected utility -->
<!--     that asset 2. -->
